---
title: "FPR on mocks - PBMC 10X-genomics - Influence of epsilon"
author: "Fanny Perraudeau"
date: "`r Sys.Date()`"
output: 
  html_document: 
    fig_height: 7
    fig_width: 7
    toc: yes
    code_folding: hide
    toc_float: yes
---

```{r options, echo=FALSE, results="hide",mesasge=FALSE, error=FALSE, include=FALSE, autodep=TRUE}
knitr::opts_chunk$set(fig.align="center", cache=TRUE, error=FALSE, message=FALSE, warning=TRUE)
library(edgeR)
library(zinbwave)
```

We want to evaluate the impact of the ridge penalty ($\epsilon$) of ZINB-WaVE on the TPR for the mock comparisons from the PBMC 10Xgenomics dataset.

# Data
```{r data}
load('../../realdata/data/core.rda')
core
ncells = 45
nreps = 30

subsets = sapply(1:nreps, function(i){
  set.seed(i)
  c1 = sample(ncol(core), ncells, replace = FALSE)
  remaining = seq_len(ncol(core))
  remaining = remaining[!remaining %in% c1]
  c2 = sample(remaining, ncells, replace = FALSE)
  stopifnot(sum(c1 %in% c2) == 0 )
  stopifnot(sum(c2 %in% c1) == 0 )
  c(c1, c2)
})
subsets = as.data.frame(t(subsets))
```

# Influence of epsilon

## Compute ZINB-WaVE posterior probabilities

```{r dataEloop}
eLoop <- core[,as.numeric(subsets[1,])]
ngenes = sum(rowSums(assay(eLoop) > 0) < 5)
cat(sprintf("Removing %s genes with only 0's", ngenes))
eLoop <- eLoop[rowSums(assay(eLoop) > 0) > 4, ]
condition = factor(rep(c("A", "B"), each = 45))
colData(eLoop)$condition = condition
design <- model.matrix(~ condition)
epsVec = 10^seq(0,16,by=2)
```

```{r zinbesp,eval=FALSE}
library(BiocParallel)
library(doParallel)
NCORES <- 2
registerDoParallel(NCORES)
register(DoparParam())
zinbList <- lapply(epsVec, function(eps){
  zinbFit(assay(eLoop), X = design, epsilon = eps)
})
save(zinbList, file = 'zinbList_epsilon.rda')
```

```{r}
load('zinbList_epsilon.rda')
```

```{r pvals}
weightsList <- lapply(zinbList, function(x){
  computeObservationalWeights(x, assay(eLoop))
})
```

## Compute p-values

We use egdeR with ZINB-WaVE posterior probabilities as observation weights.

```{r}
pvalsList <- lapply(weightsList[1:length(epsVec)], function(w){
  d <- DGEList(assay(eLoop))
  d <- edgeR::calcNormFactors(d)
  w[is.na(w)] = 1
  d$weights <- w
  d=estimateDisp(d, design)
  fit=glmFit(d,design)
  lrt=glmWeightedF(fit, coef=2, independentFiltering = TRUE)
  lrt$table$PValue
})
```

## Biological coefficient of variation (BCV)
```{r mocksEspilonBCV}
par(mfrow=c(3,4))
myplot <- lapply(1:length(weightsList), function(i){
  d <- DGEList(assay(eLoop))
  d <- edgeR::calcNormFactors(d)
  w = weightsList[[i]]
  w[is.na(w)] = 1
  d$weights <- w
  d=estimateDisp(d, design)
  plotBCV(d, main = paste0('epsilon=', epsVec[i]), ylim = c(0,6))
})
par(mfrow=c(1,1))
```

## Posterior probabilities
For zero counts
```{r mocksEspilonWeights}
par(mfrow = c(3,4))
hh = lapply(1:length(weightsList), function(i){
   hist(weightsList[[i]][assay(eLoop)==0], main = paste0('epsilon=', epsVec[i]), ylim = c(0,5e5))
})
par(mfrow = c(1,1))
```

## FPR
```{r fdrEps10xEps}
fpr = sapply(pvalsList, function(x) mean(x <= 0.05))
print(length(fpr))
plot(log10(epsVec), fpr, main = '', type = 'o',
     xlab = expression(paste(epsilon,' (log10)')), ylab = 'PCER',cex.lab = 1.5,cex.axis= 1.5,cex.main=1.5)
abline(h = 0.05, col = 'red')
```

## Histogram of p-values.

We are expecting uniform distribution for the p-values as the 2 groups contain cells that have been randomly sampled from all the cells. Therefore, the true number of DE genes is null. When $\epsilon < 10^6$ or $\epsilon > 10^{12}$, p-values are not uniformly distributed. When $10^6 < epsilon < 10^{12}$, pvalues are uniformly distributed.

```{r fdrEps10xPval}
par(mfrow = c(3,3))
hh = lapply(1:length(epsVec), function(i){
  hist(pvalsList[[i]], main = paste0('epsilon=', epsVec[i]), ylim=c(0,2000),cex.lab = 1.5,cex.axis= 1.5,cex.main=1.5, breaks = 20)
})
par(mfrow = c(1,1))
```
## Plot paper
```{r fdrEps10x,fig.width=10,fig.height=5}
par(mar = c(4.1, 4.25, 3, 1), bty = "l")
layout(cbind(matrix(1,3,3), matrix(2:10,3,3,byrow = TRUE)))
# pcer = f(eps)
fpr = sapply(pvalsList, function(x) mean(x <= 0.05))
print(length(fpr))
plot(log10(epsVec), fpr, main = '', type = 'o',
     xlab = expression(paste(epsilon,' (log10)')), ylab = 'PCER',cex.lab = 1.5,cex.axis= 1.5,cex.main=1.5)
abline(h = 0.05, col = 'red')
mtext("a" ,at=-2, padj = -1,font=2, cex=4/3)

# hist pvalues
hh = lapply(1:length(epsVec), function(i){
  hist(pvalsList[[i]], main = paste0('epsilon=', epsVec[i]), ylim=c(0,800),cex.lab = 1.5,cex.axis= 1.5,cex.main=1.5, xlab = 'p-values',breaks = 20)
  if (i==1) mtext("b" ,at=-1,padj = -1, font=2, cex=4/3)
})
```


## Conclusion

$10^6 < \epsilon < 10^{12}$ seems to be reasonable.

# sessionInfo()
```{r}
sessionInfo()
```

