---
title: "Simulations from Trapnell dataset, fold change 2"
author: "Fanny Perraudeau & Koen Van den Berge"
date: "`r Sys.Date()`"
output:
  html_document:
    fig_height: 7
    fig_width: 7
    toc: yes
    code_folding: hide
    toc_float: yes
---


```{r options, echo=FALSE, results="hide",message=FALSE, error=FALSE, include=FALSE, autodep=TRUE, warning=FALSE}
setwd("/Users/koenvandenberge/Dropbox/phdKoen/singleCell/zinbwaveZingerGithub/zinbwaveZinger/zinbwaveSimulations/trapnell_sims_fc2/")
knitr::opts_chunk$set(fig.align="center", cache=TRUE, error=FALSE, message=FALSE, warning=TRUE)
library(zinbwave)
library(BiocParallel)
library(doParallel)
library(Biobase)
library(edgeR)
library(scales)
library(DESeq2)
library(iCOBRA) # roc
library(limma)
library(genefilter) #filtered pvalues
library(MAST)
library(RColorBrewer)
library(knitr)
library(ggplot2)
library(cowplot)
# use new simulation.
#library(zingeR)
pathToParentFolder="~/Dropbox/phdKoen/singleCell/zinbwavezingerGitHub/zinbwaveZinger/"
source(paste0(pathToParentFolder,"zingeRsimulationFunctions/simulationHelpFunctions_v7_diffInZero.R"))
```


```{r cores}
NCORES <- 2
registerDoParallel(NCORES)
register(DoparParam())
```

The goal of this document is to reproduce Figure 1 from our paper. A scRNA-seq dataset from Islam dataset is simulated using zingeR simulation framework. We evaluate the performance of different DE methods.

# Simulate scRNA-seq data

## Simulating from zingeR framework

```{r data}
load("countsTrapnellProcessed.rda")
timePoint=factor(c(rep(48,85),rep(72,64)))
paramsTrapnell = getDatasetMoMPositive(counts = countsTrapnell)
```

```{r sims}
nSamples=150
grp=as.factor(rep(0:1, each = nSamples/2)) #two-group comparison
nTags=10000 #nr of features
set.seed(11)
DEind = sample(1:nTags,floor(nTags*.1),replace=FALSE) #10% DE
fcSim=(2 + rexp(length(DEind), rate = 1/2)) #fold changes
libSizes=sample(colSums(countsTrapnell),nSamples,replace=TRUE) #library sizes
simDataTrapnell <- NBsimSingleCell(foldDiff = fcSim, ind = DEind,
                                dataset = countsTrapnell, nTags = nTags,
                                group = grp,
                                verbose = TRUE, params = paramsTrapnell,
                                lib.size = libSizes, normalizeLambda=TRUE)
simDataTrapnell$counts[1:5,1:5]

```

# Methods
## RNA-seq methods

### DESeq2
```{r DESeq2}
DESeq2 <- function(counts, group, ylim = NULL, xlim = NULL){
  colData <- data.frame(group = group)
  dse <- DESeqDataSetFromMatrix(countData=counts, colData=colData, design=~group)
  colData(dse)$group <- as.factor(colData(dse)$group)
  dse <- DESeq2::estimateSizeFactors(dse, type="poscounts")
  dse <- estimateDispersions(dse)
  dse <- nbinomWaldTest(dse, betaPrior=TRUE)
  rr <- results(dse)
  cbind(pval = rr$pvalue, padj = rr$padj)
}
```

### DESeq2 no shrink

```{r DESeq2noShrink}
DESeq2_noShrink <- function(counts, group, ylim = NULL, xlim = NULL){
  colData <- data.frame(group = group)
  dse <- DESeqDataSetFromMatrix(countData=counts, colData=colData, design=~group)
  colData(dse)$group <- as.factor(colData(dse)$group)
  dse <- DESeq2::estimateSizeFactors(dse, type="poscounts")
  dse <- estimateDispersions(dse)
  dse <- nbinomWaldTest(dse, betaPrior=FALSE)
  rr <- results(dse)
  cbind(pval = rr$pvalue, padj = rr$padj)
}
```

### DESeq2 with imputation

```{r DESeq2Impute}
DESeq2_impute <- function(counts, group, ylim = NULL, xlim = NULL){
  colData <- data.frame(group = group)
  dse <- DESeqDataSetFromMatrix(countData=counts, colData=colData, design=~group)
  colData(dse)$group <- as.factor(colData(dse)$group)
  dse <- DESeq2::estimateSizeFactors(dse, type="poscounts")
  dse <- DESeq(dse)
  #dse <- estimateDispersions(dse)
  #dse <- nbinomWaldTest(dse, betaPrior=TRUE)
  rr <- results(dse)
  cbind(pval = rr$pvalue, padj = rr$padj)
}
```

# Compare dispersion estimates

```{r islamDispFC2, warning=FALSE}
counts = simDataTrapnell$counts
myfct = list(DESeq2 = DESeq2,
             DESeq2_noShrink=DESeq2_noShrink,
             DESeq2_impute=DESeq2_impute)


ylim = c(0, 11)
xlim = c(0, 16)
res = lapply(myfct, function(fct){
  fct(counts = counts, group = grp, ylim = ylim, xlim = xlim)
})

```

```{r res}
res = lapply(res, as.data.frame)
for(k in 1:length(res)) res[[k]]$padj[is.na(res[[k]]$padj)] = 1
```

## TPR vs FDR

```{r truth}
trueDE = rep(0, nTags)
trueDE[simDataTrapnell$indDE] = 1
```


```{r trapnellROCfc2}
#all methods
pp = COBRAData(pval = as.data.frame(do.call(cbind, lapply(res, '[[', 1))),
               padj = as.data.frame(do.call(cbind, lapply(res, '[[', 2))),
                truth = data.frame(status = trueDE))
cobraperf <- calculate_performance(pp, binary_truth = "status", thrs = 0.05)
cobraplot <- prepare_data_for_plot(cobraperf)
png("~/Dropbox/phdKoen/singleCell/zinbwaveZinger/plots2/trapnellSimDESeq2.png", width=7,height=8, units="in", res=300)
plot_fdrtprcurve(cobraplot, pointsize=1) + xlab("FDP")
dev.off()

```
