---
title: "Compare DE methods for 10X genomics"
author: "Fanny Perraudeau"
date: "`r Sys.Date()`"
output: 
  html_document: 
  fig_height: 7
fig_width: 7
toc: yes
code_folding: hide
toc_float: yes
---
  
```{r options, echo=FALSE, results="hide",mesasge=FALSE, error=FALSE, include=FALSE, autodep=TRUE}
knitr::opts_chunk$set(fig.align="center", cache=TRUE, error=FALSE, message=FALSE, warning=TRUE)
library(zinbwave)
library(edgeR)
library(RColorBrewer)
library(ggplot2)
```

Using our clustering, we want to compare the different DE methods and perform gene set enrichment analysis.

# Data
```{r loadobject}
load("coreAfterClustering.rda")
dim(core)
core = core[, colData(core)$ours %in% c(0,5)]
dim(core)
keepCells = colnames(core)
temp = core

load('../data/core.rda')
dim(core)
core = core[, keepCells]
dim(core)
core = core[rowSums(assay(core)) > 0, ]
dim(core)
colData(core)$seurat = factor(colData(core)$seurat)
ours = colData(temp)$ours
nn = gsub('-1', '', names(ours))
ours = as.integer(as.vector(ours))
names(ours) = nn
colData(core)$ours = factor(ours[colnames(core)])
```

# Compute ZINB-WaVE posterior probabilities

```{r zinb,eval=FALSE}
# also tried eps = 10e8, 10e10
library(doParallel)
library(BiocParallel)
NCORES = 2
registerDoParallel(NCORES)
register(DoparParam())
print(system.time(zinb <- zinbFit(core, X = '~ ours',
                                  epsilon = 1e12)))
save(zinb, file = 'zinbX.rda')
```

```{r weightszinb}
load('zinbX.rda')
counts = assay(core)
weights_zinbwave = computeObservationalWeights(zinb, counts)
hist(weights_zinbwave, main = 'zinbwave', xlab = 'Weights')
```

# DE methods

```{r}
counts = assay(core)
design = model.matrix(~ colData(core)$ours)
```

## Run edgeR
```{r fitedger}
fit_edgeR <- function(counts, design, filter = NULL){
  library(edgeR)
  d = DGEList(counts)
  d = suppressWarnings(calcNormFactors(d))
  d = estimateDisp(d, design)
  fit = glmFit(d, design)
  glm = glmLRT(fit)
  tab = glm$table
  tab$padj = p.adjust(tab$PValue, "BH")
  tab$gene = rownames(tab)
  de <- as.data.frame(tab, stringsAsFactors = FALSE)
  de = de[, c('gene', 'PValue', 'padj', 'logFC')]
  colnames(de) = c('gene', 'pval', 'padj', 'logfc')
  de
}
```

```{r edgeR}
edgeR <- fit_edgeR(counts, design)
edgeR$method <- 'edgeR'
```

## Run edgeR with zinbwave weights
```{r fitedgeRzi}
fit_edgeR_zi <- function(counts, design, weights, filter = NULL){
  library(edgeR)
  d = DGEList(counts)
  d = suppressWarnings(calcNormFactors(d))
  d$weights <- weights 
  d = estimateDisp(d, design)
  fit = glmFit(d,design)
  glm = glmWeightedF(fit, filter = filter)
  tab = glm$table
  tab$gene = rownames(tab)
  de <- data.frame(tab, stringsAsFactors = FALSE)
  de = de[, c('gene', 'PValue', 'padjFilter', 'logFC')]
  colnames(de) = c('gene', 'pval', 'padj', 'logfc')
  de
}
```

```{r zinbwaveedgeR}
# check if filter ok with baseMean, not sure
# filter just affect adjusted pvalues and we don't use 
# much pvalues
nf <- edgeR::calcNormFactors(counts)
baseMean = unname(rowMeans(sweep(counts,2,nf,FUN="*")))
zinbwave_edgeR <- fit_edgeR_zi(counts, design, 
                               weights = weights_zinbwave,
                               filter = baseMean)
zinbwave_edgeR$method <- 'zinbwave_edgeR'
```

## Run Limma-voom
```{r runlimmavoom}
runLimmavoom <- function(counts, design) {
  library(limma)
  library(edgeR)
  dgel <- DGEList(counts)
  dgel <- edgeR::calcNormFactors(dgel)
  v <- voom(dgel,design,plot=FALSE)
  fit <- lmFit(v,design)
  fit <- eBayes(fit)
  tt <- topTable(fit,coef=2,n=nrow(dgel),sort.by="none")
  pvals <- tt$P.Value
  padj <- p.adjust(pvals,method="BH")
  padj[is.na(padj)] <- 1
  data.frame(gene = rownames(tt), pval=pvals, padj=padj, logfc=tt$logFC)
}
```

```{r limmavoom}
voom <- runLimmavoom(counts, design)
voom$method <- 'limmavoom'
```

## Run MAST
```{r runMAST}
runMAST <- function(counts, design) {
  library(MAST)
  tpm <- counts*1e6/colSums(counts)
  tpm <- log2(tpm+1)
  sca <- FromMatrix(tpm,cData=data.frame(group=factor(design[,2])))
  # here, we keep all genes so that we can fairly compare MAST and the other methods. So, no adaptive thresholding or filtering by gene expression
  assays(sca) <- list(tpm=assay(sca))
  ngeneson <- apply(counts,2,function(x) mean(x>0))
  CD <- colData(sca)
  CD$ngeneson <- ngeneson
  CD$cngeneson <- CD$ngeneson-mean(ngeneson)
  colData(sca) <- CD
  ## differential expression
  fit <- zlm(~ cngeneson + group, sca = sca,
             method = "bayesglm", ebayes = TRUE)
  summaryDt = summary(fit, doLRT='group1')
  summaryDt = summaryDt$datatable
  fcHurdle <- merge(summaryDt[contrast=='group1'&component=='H',.(primerid, `Pr(>Chisq)`)],
                    summaryDt[contrast=='group1' & component=='logFC', .(primerid, coef, ci.hi, ci.lo)], by='primerid')
  fcHurdle[,padj:=p.adjust(`Pr(>Chisq)`, 'fdr')]
  df = data.frame(gene = fcHurdle$primerid,
             pval=fcHurdle[,'Pr(>Chisq)'], padj=fcHurdle$padj,
             logfc=fcHurdle$coef)
  colnames(df)[2] = 'pval'
  df
}
```

```{r mast}
mast <- runMAST(counts, design)
mast$method <- 'MAST'
```

## Load Seurat results
```{r}
seurat = read.csv('tcellmarkers_seurat.csv')
seurat$padj = NA
seurat$method = 'seurat'
seurat = seurat[, c('gene', 'p_val', 'p_val_adj',
                    'avg_logFC', 'method')]
colnames(seurat) = colnames(edgeR)
```

## Concatenate all the result tables
```{r}
de = rbind(edgeR, seurat, voom, zinbwave_edgeR, mast)
write.csv(de, file = 'deGenes.csv', row.names = FALSE)
```

```{r}
de = read.csv('deGenes.csv', stringsAsFactors = FALSE)
head(de,2)
```

# Compare methods

## GSEA pre ranked
```{r}
library(xCell) # for db
library(fgsea)
library(GSEABase)

## extract genesets from xcell
nagenes = unique(de[is.na(de$logfc), 'gene'])
de = de[!de$gene %in% nagenes, ]
genesets <- unlist(geneIds(xCell.data$signatures))
celltypes <- sapply(strsplit(names(genesets), "%"), function(x) x[1])
names(genesets) <- NULL
gs <- tapply(genesets, celltypes, c)
set.seed(6372)
gsea_res = lapply(unique(de$method), function(x){
  print(x)
  temp = de[de$method == x, ]
  pval = temp$pval
  zscores = qnorm(1 - (pval/2))
  zscores[is.infinite(zscores)] = max(zscores[!is.infinite(zscores)])
  logfc = temp$logfc
  zscores[logfc<0] = -zscores[logfc<0]
  names(zscores) = temp$gene
  if (x == 'seurat') zscores = -zscores
  gsea = fgsea(gs, zscores, nperm = 10000, minSize = 5)
  gsea$method = x
  gsea[order(-abs(gsea$NES)), ]
})
lapply(gsea_res, head)
```

```{r}
gseaDf = as.data.frame(do.call(rbind, gsea_res))
gseaDf = gseaDf[gseaDf$size > 100, ]
gseaDf = gseaDf[, c('method', 'pathway', 'NES')]
#gseaDf$method = factor(gseaDf$method, levels = c('edgeR', 'seurat', 'MAST', 'zinbwave_DESeq2', 'limmavoom', 'zinbwave_edgeR'))
sortedPwy = gseaDf[gseaDf$method == 'zinbwave_edgeR', ]
sortedPwy = sortedPwy[order(sortedPwy$NES), 'pathway']
gseaDf$pathway = factor(gseaDf$pathway, levels = sortedPwy)
```

```{r}
write.csv(gseaDf, file = 'gsea.csv')
```

```{r tenxcaseNESallW}
size=20
ggplot(gseaDf, aes(method, pathway)) +
  geom_tile(aes(fill = NES)) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_fill_gradient2(low = "blue", high = "red", mid = "white", 
                       midpoint = 0, space = "Lab", 
                       name="Normalized\nEnrichment\nScore") +
  ylab('Cell Type') + xlab('Method') 
```

```{r tenxcaseNEScd4W}
chosen = c('CD4+ memory T-cells','CD4+ Tem','CD4+ Tcm','CD4+ naive T-cells')
sub = gseaDf[gseaDf$pathway %in% chosen, ]
nes = sub$NES
limit = max(abs(min(nes)), max(nes))
sub$pathway = factor(sub$pathway, levels = chosen)
ggplot(sub, aes(method, pathway)) +
  geom_tile(aes(fill = NES)) + 
  scale_fill_gradient2(low = "blue", high = "red", 
                       mid = "white", 
                       midpoint = 0, space = "Lab", name="Normalized\nEnrichment\nScore", 
                       limit = c(-limit, limit)) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  ylab('Cell Type') + xlab('Method') +
  theme(text = element_text(size = size))
```

```{r}
groups = grep('CD4+', gsea_res[[1]]$pathway, value =T)
lapply(gsea_res, function(x){
  x[x$pathway %in% groups, ]
})
```


## Histogram of pvalues

We are expecting uniformity of the pvalue with a pick close to zero corresponding to DE genes.

```{r tenxcaseHistPvalW}
par(mfrow= c(3,2))
for (x in unique(de$method)){
  hist(de[de$method == x, 'pval'], main = x, ylim = c(0, 3000),
       xlab = 'pvalue')
}
par(mfrow= c(1,1))
```
## Number of DE genes
```{r}
for (x in unique(de$method)){
  print(x)
  print(sum(de[de$method ==x, 'padj'] < 0.05, na.rm = TRUE))
}
```

## Concordance between DE genes
### Venn diagram

```{r}
thr = 0.05
```

DE genes are genes with an adjusted pvalue lower than `r thr`.

```{r venn,eval=FALSE}
library(dplyr)
ve = de %>% group_by(method, gene) %>%
  summarize(pval = min(pval, na.rm=TRUE)) %>%
  ungroup() %>% as.data.frame()
ve$de = ve$pval < thr
ve = reshape2::acast(ve[, c(1,2,4)], gene ~ method, sum)

aa <- vennCounts(ve)
vennDiagram(aa, main = 'DE gene, adj pvalue < 0.05')
```


## Most DE genes per method in terms of pvalue
```{r}
ranks = lapply(unique(de$method), function(x){
  temp = de[de$method == x, ]
  temp$rank = rank(temp$pval)
  temp
})
de = do.call(rbind, ranks)

ranks = lapply(unique(de$method), function(x){
  temp = de[de$method == x, ]
  temp = temp[temp$rank < 10, ]
  temp[order(temp$rank), ]
})
names(ranks) = unique(de$method)
ranks
```

## Most DE genes per method in terms of fold change
```{r}
ranks = lapply(unique(de$method), function(x){
  temp = de[de$method == x, ]
  temp$rank = rank(-abs(temp$logfc))
  temp
})
de = do.call(rbind, ranks)

ranks = lapply(unique(de$method), function(x){
  temp = de[de$method == x, ]
  temp = temp[temp$rank < 10, ]
  temp[order(temp$rank), ]
})
names(ranks) = unique(de$method)
ranks
```

## Heatmaps 

```{r tcnorm}
norm10x <- function (ei){
  sums = colSums(ei)
  t(t(ei)*median(sums)/sums)
}
tc = norm10x(counts) 
tc = log1p(tc)
```

```{r heatmaptc}
library(clusterExperiment)
tt = lapply(unique(de$method), function(x){
  temp = de[de$method == x, ]
  degenes = temp[order(temp[, 'pval'])[1:10], 'gene']
  cc = as.numeric(colData(core)$ours)
  plotHeatmap(tc[rownames(tc) %in% degenes, ], main = x,
              breaks = .99,
              sampleData = data.frame(clusters = cc))
})
```

## Visualize clustering with DE genes

DE genes are genes with a pvalue lower than `r thr`. We subset the data to keep only DE genes and visualize the clusters in 2D.

```{r pcaDE}
library(rARPACK) 
library(Rtsne)
fastpca <- function(expr, scale=FALSE, k = 25) {
  svd_raw <- svds(scale(t(expr), center=TRUE, scale=scale),
                  k=k, nu=k, nv=0)
  pc_raw <- svd_raw$u %*% diag(svd_raw$d[1:k])
  return(pc_raw)
}

par(mfrow = c(2,2))
for (x in unique(de$method)){
  temp = de[de$method == x, ]
  degenes = temp[temp$pval< 0.05, 'gene']
  ngenes = length(degenes)
  pca = fastpca(tc[rownames(tc) %in% degenes, ])
  pca_tsne <- Rtsne(pca, pca = FALSE)
  plot(pca_tsne$Y, pch=19, col=as.numeric(colData(core)$ours),
       main=paste(x, ', ', ngenes, ' DE genes'))
}
par(mfrow = c(1,1))
```


#sessionInfo
```{r}
sessionInfo()
```
