---
title: "Dimensionality reduction using zinbwave - 10x genomics"
author: "Fanny Perraudeau"
date: "`r Sys.Date()`"
output: 
html_document: 
fig_height: 7
fig_width: 7
toc: yes
code_folding: hide
toc_float: yes
---

```{r options, echo=FALSE, results="hide",mesasge=FALSE, error=FALSE, include=FALSE, autodep=TRUE}
knitr::opts_chunk$set(fig.align="center", cache=TRUE, error=FALSE, message=FALSE, warning=TRUE)
library(zinbwave)
library(edgeR)
library(zingeR)
library(RColorBrewer)
library(Rtsne)
library(ggplot2)
set.seed(7383)
```

After running Seurat preprocessing, we want to perform dimensionality reduction using ZINB-WaVE.

# Data 
```{r loadobject}
load("../core.rda")
core = core[rowSums(assay(core)) > 0, ]
core
colData(core)$seurat = factor(colData(core)$seurat)
```

Select only most variable genes used for Seurat clustering
```{r keepgenes}
core = core[rowData(core)$seuratVarGenes, ]
dim(core)
```

# Chosen K in zinbwave

We chose K in zinbwave using AIC and BIC.

```{r aicbic}
library(doParallel)
library(BiocParallel)
NCORES <- 2
registerDoParallel(NCORES)
register(DoparParam())

kvec = c(0, 3, 5, 8, 10, 15, 20, 25, 30, 45)
if (FALSE){
  for (k in kvec){
    print(k)
    #tried eps = 1e4
    print(system.time(zinb <- zinbFit(core, epsilon = 1e12, K = k)))
    fn = sprintf('zinb_k%s.rda', k)
    save(zinb, file = fn)
  }
}

# compute aic, bic 
# these functions should be available in zinbwave package soon
zinbAIC <- function(model, x) {
  if ((nSamples(model) != nrow(x))|(nFeatures(model) != ncol(x))) {
    stop("x and model should have the same dimensions!")
  }
  k <- nParams(model)
  ll <- loglik(model, x)
  return(2*k - 2*ll)
}

zinbBIC <- function(model, x) {
  n <- nSamples(model)
  if ((n != nrow(x))|(nFeatures(model) != ncol(x))) {
    stop("x and model should have the same dimensions!")
  }
  k <- nParams(model)
  ll <- loglik(model, x)
  return(log(n)*k - 2*ll)
}

res = sapply(kvec, function(k){
  fn = sprintf('zinb_k%s.rda', k)
  load(fn)
  aic = zinbAIC(zinb, t(assay(core)))
  bic = zinbBIC(zinb, t(assay(core)))
  c(aic = aic, bic = bic)
})
```

```{r tenxcaseAICBIC}
par(mfrow=c(1,2))
plot(kvec, res['aic', ], type = 'o', main = 'AIC', xlab = 'K', ylab = 'AIC')
plot(kvec, res['bic', ], type = 'o', main = 'BIC', xlab = 'K', ylab = 'BIC')
par(mfrow=c(1,1))
```

From the previous plots, we decide to use K = 20 as suggested by AIC.

```{r loadChosen}
load('zinb_k20.rda')
```

# TSNE on W

Let's visualize W using tSNE. Colors are from labels found by Seurat pipeline.

```{r plotK}
W = getW(zinb)
w_tsne <- Rtsne(W, pca = FALSE, perplexity = 200)
zinb_tsne <- as.data.frame(w_tsne$Y)
colnames(zinb_tsne) = c('tSNE_1', 'tSNE_2')
ggplot(zinb_tsne, aes(x = tSNE_1, y = tSNE_2)) + 
  geom_point(mapping = aes(colour = colData(core)$seurat)) + 
  theme_bw() + 
  theme(legend.title = element_blank(),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank())
```


# sessionInfo
```{r}
sessionInfo()
```