---
  title: "DE using ZINB-WaVE observational weights - Usoskin dataset"
author: "Fanny Perraudeau"
date: "`r Sys.Date()`"
output:
  html_document:
  fig_height: 7
fig_width: 7
toc: yes
code_folding: hide
toc_float: yes
---

  ```{r options, echo=FALSE, results="hide",message=FALSE, error=FALSE, include=FALSE, autodep=TRUE}
knitr::opts_chunk$set(fig.align="center", cache=TRUE, error=FALSE, message=FALSE, warning=TRUE)
library(SingleCellExperiment)
```

We want to find the DE genes between the cell types. For clustering, we used only the 1,000 most variable genes. Here, we use all the genes.

# Data
```{r}
load('seAfterClustering.rda')
dim(se)
table(colData(se)$ourClusters, colData(se)$Level3)
se = se[, colData(se)$Level3 == 'TH']
dim(se)
colData(se)$ourClusters[colData(se)$ourClusters=='c2'] = 'TH1'
colData(se)$ourClusters[colData(se)$ourClusters=='c7'] = 'TH2'
table(colData(se)$ourClusters, colData(se)$Pickingsessions)
table(colData(se)$ourClusters, colData(se)$Sex)
table(colData(se)$Pickingsessions, colData(se)$Sex)
```


```{r}
load('data/esetUsoskin.RData')
eset=eset[rowSums(exprs(eset)>0)>=20,]
exprs(eset) <- apply(exprs(eset),2,function(x) {storage.mode(x) <- 'integer'; x})
cData = pData(eset)
colnames(cData) = gsub(' ', '', colnames(cData))
seAll = SingleCellExperiment(assays = list(counts = exprs(eset)),
                             colData = cData,
                             rowData = fData(eset))
colData(seAll)$Pickingsessions = factor(colData(seAll)$Pickingsessions)
pal <- clusterExperiment::bigPalette
colData(seAll)$Level3 <- as.vector(colData(seAll)$Level3)
pal = pal[1:length(unique(colData(seAll)$Level3))]
names(pal) = unique(colData(seAll)$Level3)
dim(seAll)
seAll=seAll[, colnames(seAll) %in% colnames(se)]
seAll = seAll[,colnames(se)]
colData(seAll) = colData(se)
se = seAll
se = se[rowSums(assay(se)) > 0, ]
dim(se)
```

# ZINB-WaVE observational weights

```{r}
# This function should be soon in zinbwave package
computeObservationalWeights <- function(model, x){
  mu <- getMu(model)
  pi <- getPi(model)
  theta <- getTheta(model)
  theta <- matrix(rep(theta, each = ncol(x)), ncol = nrow(x))
  nb_part <- dnbinom(t(x), size = theta, mu = mu)
  zinb_part <- pi * ( t(x) == 0 ) + (1 - pi) *  nb_part
  zinbwg <- ( (1 - pi) * nb_part ) / zinb_part
  t(zinbwg)
}
```

```{r,eval=FALSE}
library(zinbwave)
library(doParallel)
library(BiocParallel)
NCORES <- 2
registerDoParallel(NCORES)
register(DoparParam())
set.seed(8492)

print(system.time(zinb <- zinbFit(se, epsilon = 1e8, X = '~ Pickingsessions + ourClusters')))
fn = 'zinb_batch.rda'
save(zinb, file = fn)
```

```{r}
load('zinb_batch.rda')
weights_batch = computeObservationalWeights(zinb, assay(se))
hist(weights_batch)
```

# Run DE methods

```{r}
weights_zinbwave = weights_batch
colData(se)$ourClusters = factor(colData(se)$ourClusters)
colData(se)$Pickingsessions = factor(colData(se)$Pickingsessions)
design = model.matrix(~ colData(se)$Pickingsessions +
                        colData(se)$ourClusters)
counts = assay(se)
rownames(counts) = rowData(se)[,1]
```

# Run edgeR
```{r fitedger}
fit_edgeR <- function(counts, design, filter = NULL){
  library(edgeR)
  d = DGEList(counts)
  d = suppressWarnings(calcNormFactors(d))
  d = estimateDisp(d, design)
  fit = glmFit(d, design)
  glm = glmLRT(fit)
  tab = glm$table
  tab$padj = p.adjust(tab$PValue, "BH")
  tab$gene = rownames(tab)
  de <- as.data.frame(tab, stringsAsFactors = FALSE)
  de = de[, c('gene', 'PValue', 'padj', 'logFC')]
  colnames(de) = c('gene', 'pval', 'padj', 'logfc')
  de
}
```

```{r edgeR}
edgeR <- fit_edgeR(counts, design)
edgeR$method <- 'edgeR'
```

# Run edgeR with zinbwave weights
```{r fitedgeRzi}
fit_edgeR_zi <- function(counts, design, weights,
                         filter = NULL){
  library(edgeR)
  library(zingeR)
  d = DGEList(counts)
  d = suppressWarnings(calcNormFactors(d))
  d$weights <- weights
  d = estimateDisp(d, design)
  fit = glmFit(d,design)
  glm = glmWeightedF(fit, filter = filter)
  tab = glm$table
  tab$gene = rownames(tab)
  de <- data.frame(tab, stringsAsFactors = FALSE)
  de = de[, c('gene', 'PValue', 'padjFilter', 'logFC')]
  colnames(de) = c('gene', 'pval', 'padj', 'logfc')
  de
}
```

```{r zinbwaveedgeR}
# check if filter ok with baseMean, not sure
# filter just affect adjusted pvalues and we don't use
# much pvalues
nf <- edgeR::calcNormFactors(counts)
baseMean = unname(rowMeans(sweep(counts,2,nf,FUN="*")))
zinbwave_edgeR <- fit_edgeR_zi(counts, design,
                               weights = weights_zinbwave,
                               filter = baseMean)
zinbwave_edgeR$method <- 'zinbwave_edgeR'
```


# Run DESeq2 with zinbwave weights
```{r fitDESeq2zi}
fit_DESeq2_zi <- function(counts, design, weights){
  library(DESeq2)
  cData = data.frame(ourClusters = factor(design[,4]))
  Pickingsessions = rep('Cold', nrow(cData))
  Pickingsessions[design[, 2] == 1] = 'RT1'
  Pickingsessions[design[, 3] == 1] = 'RT2'
  cData$Pickingsessions = Pickingsessions
  cData$Pickingsessions=factor(cData$Pickingsessions)
  cData$ourClusters=factor(cData$ourClusters)
  rownames(cData) = colnames(counts)
  dds <- DESeqDataSetFromMatrix(counts, colData =cData, design = ~ Pickingsessions + ourClusters)
  dds$design <- design
  assays(dds)[["weights"]] <- weights
  dds <- estimateSizeFactors(dds, type = "poscounts")
  dds <- estimateDispersions(dds)
  dds <- nbinomWaldTest(dds, betaPrior = TRUE, useT = TRUE,
                        df = rowSums(weights) - 2)
  tab <- results(dds)
  tab$gene = rownames(tab)
  de <- data.frame(tab, stringsAsFactors = FALSE)
  de = de[, c('gene', 'pvalue', 'padj', 'log2FoldChange')]
  colnames(de) = c('gene', 'pval', 'padj', 'logfc')
  de
}
```

```{r zinbwavedeseq2}
zinbwave_deseq2 <- fit_DESeq2_zi(counts, design,
                                 weights = weights_zinbwave)
zinbwave_deseq2$method <- 'zinbwave_DESeq2'
```

# Run DESeq2
```{r fitDESeq2}
fit_DESeq2 <- function(counts, design){
  library(DESeq2)
  cData = data.frame(ourClusters = factor(design[,4]))
  Pickingsessions = rep('Cold', nrow(cData))
  Pickingsessions[design[, 2] == 1] = 'RT1'
  Pickingsessions[design[, 3] == 1] = 'RT2'
  cData$Pickingsessions = Pickingsessions
  cData$Pickingsessions=factor(cData$Pickingsessions)
  cData$ourClusters=factor(cData$ourClusters)
  rownames(cData) = colnames(counts)
  dds <- DESeqDataSetFromMatrix(counts, colData =cData, design = ~ Pickingsessions + ourClusters)
  dds$design <- design
  dds <- estimateSizeFactors(dds, type = "poscounts")
  dds <- estimateDispersions(dds)
  dds <- nbinomWaldTest(dds, betaPrior = TRUE)
  tab <- results(dds)
  tab$gene = rownames(tab)
  de <- data.frame(tab, stringsAsFactors = FALSE)
  de = de[, c('gene', 'pvalue', 'padj', 'log2FoldChange')]
  colnames(de) = c('gene', 'pval', 'padj', 'logfc')
  de
}
```

```{r deseq2}
deseq2 <- fit_DESeq2(counts, design)
deseq2$method <- 'DESeq2'
```

# Run MAST
```{r}
runMAST <- function(counts, design){
  require(MAST)
  tpm <- counts*1e6/colSums(counts)
  tpm <- log2(tpm+1)
  sca <- FromMatrix(tpm, cData=data.frame(group=design[,4]))

  # Adaptive thresholding from MAST vignette
  thres <- thresholdSCRNACountMatrix(assay(sca), nbins = 50, min_per_bin = 50,
                                     conditions = design[,4])
  #par(mfrow=c(5,4))
  #plot(thres)
  assays(sca) <- list(thresh=thres$counts_threshold, tpm=assay(sca))

  ngeneson <- apply(counts,2,function(x) mean(x>0))

  # include gene detection prop in the covariates
  CD <- colData(sca)
  CD$ngeneson <- ngeneson
  CD$cngeneson <- CD$ngeneson-mean(ngeneson)
  colData(sca) <- CD

  ## differential expression
  fit <- zlm(~ cngeneson + group, sca = sca, method = "bayesglm", ebayes = TRUE)
  summaryCond <- summary(fit, doLRT = 'group')
  summaryDt <- summaryCond$datatable
  tab <- merge(summaryDt[contrast=='group' & component=='H',
                         .(primerid, `Pr(>Chisq)`)],
               summaryDt[contrast=='group' & component=='logFC',
                         .(primerid, coef)], by='primerid')
  pval <- tab[, `Pr(>Chisq)`]
  padj <- p.adjust(pval, method = "BH")
  data.frame(gene = tab$primerid, pval= pval, padj = padj, logfc = tab$coef)
}
```

```{r mast}
mast <- runMAST(counts, design)
mast$method <- 'MAST'
```

# Run Limma-voom
```{r runlimmavoom}
runLimmavoom <- function(counts, design) {
  library(limma)
  library(edgeR)
  dgel <- DGEList(counts)
  dgel <- edgeR::calcNormFactors(dgel)
  v <- voom(dgel,design,plot=FALSE)
  fit <- lmFit(v,design)
  fit <- eBayes(fit)
  tt <- topTable(fit,coef=2,n=nrow(dgel),sort.by="none")
  pvals <- tt$P.Value
  padj <- p.adjust(pvals,method="BH")
  padj[is.na(padj)] <- 1
  data.frame(gene = rownames(tt), pval=pvals, padj=padj, logfc=tt$logFC)
}
```

```{r limmavoom}
voom <- runLimmavoom(counts, design)
voom$method <- 'limmavoom'
```

# Concatenate all the result tables
```{r}
# todo: add deseq2
de = rbind(edgeR, voom, zinbwave_edgeR)
write.csv(de, file = 'deGenes.csv', row.names = FALSE)
```

```{r}
de = read.csv('deGenes.csv', stringsAsFactors = FALSE)
head(de,2)
```

# Compare methods

## GSEA pre ranked
```{r}
library(xCell) # for db
library(fgsea)
library(GSEABase)

## extract genesets from xcell
genesets <- unlist(geneIds(xCell.data$signatures))
celltypes <- sapply(strsplit(names(genesets), "%"), function(x) x[1])
names(genesets) <- NULL
gs <- tapply(genesets, celltypes, c)
de$logfc[is.na(de$logfc)] = rnorm(n = sum(is.na(de$logfc)), 0, 0.005)
# MAST creates NAs for log fold change but fgsea does not handle NAs, so I artificially creates small log fold change. rnorm with mean = 0, and not 0 because fgsea does not handle tights.
set.seed(6372)
gsea_res = lapply(unique(de$method), function(x){
  print(x)
  temp = de[de$method == x, ]
  degenes = temp$logfc
  names(degenes) = toupper(temp$gene)
  gsea = fgsea(gs, degenes, nperm = 1000, minSize = 10)
  gsea$method = x
  gsea$NES[is.na(gsea$NES)] = rnorm(n = sum(is.na(gsea$NES)), 0, 0.005)
  gsea[order(-abs(gsea$NES)), ]
})
lapply(gsea_res, head)
```

```{r}
gseaDf = as.data.frame(do.call(rbind, gsea_res))
gseaDf = gseaDf[, c('method', 'pathway', 'NES')]
gseaDf$method = factor(gseaDf$method, levels = c('limmavoom','edgeR',  'zinbwave_edgeR'))
sortedPwy = gseaDf[gseaDf$method == 'zinbwave_edgeR', ]
sortedPwy = sortedPwy[order(sortedPwy$NES), 'pathway']
gseaDf$pathway = factor(gseaDf$pathway, levels = sortedPwy)
```

```{r}
library(ggplot2)
ggplot(gseaDf, aes(method, pathway)) +
  geom_tile(aes(fill = NES)) +
  scale_fill_gradient2(low = "blue", high = "red", mid = "white",
                       midpoint = 0, space = "Lab",
                       name="Enrichment score")
```

## Gene ontology

```{r,eval=FALSE}
fl <- "http://www.geneontology.org/ontology/subsets/goslim_generic.obo"
fl <- "http://data.pantherdb.org/PANTHER12/ontology/PANTHERGOslim.obo"
obo = getOBOCollection(fl, evidenceCode="TAS")

subsets(obo)
obo['goslim_mouse']


rownames(se) = rowData(se)[,1]
es=as(se, "ExpressionSet")
annotation(es)="hgu95av2"
gsc <- GeneSetCollection(es, setType = GOCollection())

GeneSetCollection(rowData(se)[,1], idType = , setType="GOCollection")

## Not run:


genesets <- unlist(geneIds(xCell.data$signatures))
celltypes <- sapply(strsplit(names(genesets), "%"), function(x) x[1])
names(genesets) <- NULL
gs <- tapply(genesets, celltypes, c)
de$logfc[is.na(de$logfc)] = rnorm(n = sum(is.na(de$logfc)), 0, 0.005)
# MAST creates NAs for log fold change but fgsea does not handle NAs, so I artificially creates small log fold change. rnorm with mean = 0, and not 0 because fgsea does not handle tights.
set.seed(6372)
gsea_res = lapply(unique(de$method), function(x){
  print(x)
  temp = de[de$method == x, ]
  degenes = temp$logfc
  names(degenes) = toupper(temp$gene)
  gsea = fgsea(gs, degenes, nperm = 1000, minSize = 10)
  gsea$method = x
  gsea$NES[is.na(gsea$NES)] = rnorm(n = sum(is.na(gsea$NES)), 0, 0.005)
  gsea[order(-abs(gsea$NES)), ]
})
lapply(gsea_res, head)
```

```{r,eval=FALSE}}
temp = de[de$method == 'zinbwave_edgeR', ]
degenes = temp[abs(temp$logfc) > 1 & temp$pval < 0.05, 'gene']
write.csv(data.frame(genes = degenes),'de_zinbwave.csv')

temp = de[de$method == 'edgeR', ]
degenes = temp[abs(temp$logfc) > 1 & temp$pval < 0.05, 'gene']
write.csv(data.frame(genes = degenes),'de_edger.csv')

temp = de[de$method == 'limmavoom', ]
degenes = temp[abs(temp$logfc) > 1 & temp$pval < 0.05, 'gene']
write.csv(data.frame(genes = degenes),'de_limma.csv')
```




## Histogram of pvalues

We are expecting uniformity of the pvalue with a pick close to zero corresponding to DE genes.

```{r hist}
par(mfrow= c(3,3))
for (x in unique(de$method)){
  hist(de[de$method == x, 'pval'], main = x, ylim = c(0, 3000),
       xlab = 'pvalue')
}
par(mfrow= c(1,1))
```


## Concordance between DE genes
### Venn diagram

```{r}
thr = 0.05
```

DE genes are genes with an adjusted pvalue lower than `r thr`.

```{r venn,eval=FALSE}
library(dplyr)
ve = de %>% group_by(method, gene) %>%
  summarize(pval = min(pval, na.rm=TRUE)) %>%
  ungroup() %>% as.data.frame()
ve$de = ve$pval < thr
ve = reshape2::acast(ve[, c(1,2,4)], gene ~ method, sum)

aa <- vennCounts(ve)
vennDiagram(aa, main = 'DE gene, adj pvalue < 0.05')
```


## Most DE genes per method in terms of pvalue
```{r}
ranks = lapply(unique(de$method), function(x){
  temp = de[de$method == x, ]
  temp$rank = rank(temp$pval)
  temp
})
de = do.call(rbind, ranks)

ranks = lapply(unique(de$method), function(x){
  temp = de[de$method == x, ]
  temp = temp[temp$rank < 10, ]
  temp[order(temp$rank), ]
})
names(ranks) = unique(de$method)
ranks
```

## Most DE genes per method in terms of fold change
```{r}
ranks = lapply(unique(de$method), function(x){
  temp = de[de$method == x, ]
  temp$rank = rank(-abs(temp$logfc))
  temp
})
de = do.call(rbind, ranks)

ranks = lapply(unique(de$method), function(x){
  temp = de[de$method == x, ]
  temp = temp[temp$rank < 10, ]
  temp[order(temp$rank), ]
})
names(ranks) = unique(de$method)
ranks
```

## Heatmaps

```{r tcnorm}
norm10x <- function (ei){
  sums = colSums(ei)
  t(t(ei)*median(sums)/sums)
}
tc = norm10x(counts)
tc = log1p(tc)
```

```{r heatmaptc}
library(clusterExperiment)
tt = lapply(unique(de$method), function(x){
  temp = de[de$method == x, ]
  degenes = temp[order(temp[, 'pval'])[1:50], 'gene']
  cc = as.numeric(colData(se)$ourClusters)
  plotHeatmap(tc[rownames(tc) %in% degenes, ], main = x,
              breaks = .98,
              sampleData = data.frame(clusters = cc))
})
```

## Visualize clustering with DE genes

DE genes are genes with a pvalue lower than `r thr`. We subset the data to keep only DE genes and visualize the clusters in 2D.

```{r pcaDE}
library(rARPACK)
library(Rtsne)
fastpca <- function(expr, scale=FALSE, k = 25) {
  svd_raw <- svds(scale(t(expr), center=TRUE, scale=scale),
                  k=k, nu=k, nv=0)
  pc_raw <- svd_raw$u %*% diag(svd_raw$d[1:k])
  return(pc_raw)
}

par(mfrow = c(3,3))
for (x in unique(de$method)){
  temp = de[de$method == x, ]
  degenes = temp[temp$pval < 0.05, 'gene']
  ngenes = length(degenes)
  pca = fastpca(tc[rownames(tc) %in% degenes, ])
  pca_tsne <- Rtsne(pca, pca = FALSE)
  plot(pca_tsne$Y, pch=19, col=as.numeric(colData(se)$ourClusters),
       main=paste(x, ', ', ngenes, ' DE genes'))
}
par(mfrow = c(1,1))
```


