---
title: "Usoskin DE analysis"
author: "Koen Van den Berge"
date: "`r Sys.Date()`"
output:
  html_document:
    fig_height: 7
    fig_width: 7
    toc: yes
    code_folding: hide
    toc_float: yes
---

This document uses Fanny's original simulation code in `islam_sims_fc2.Rmd' and has been adapted by Koen to adopt the new simulation framework.

```{r options, echo=FALSE, results="hide",message=FALSE, error=FALSE, include=FALSE, autodep=TRUE, warning=FALSE}
setwd("/Users/koenvandenberge/Dropbox/phdKoen/singleCell/zinbwaveZingerGithub/zinbwaveZinger/realdata/usoskin/")
knitr::opts_chunk$set(fig.align="center", cache=TRUE, error=FALSE, message=FALSE, warning=TRUE)
library(zinbwave)
library(BiocParallel)
library(doParallel)
library(Biobase)
library(edgeR)
library(scales)
library(DESeq2)
library(iCOBRA) # roc
library(limma)
library(genefilter) #filtered pvalues
library(MAST)
library(RColorBrewer)
library(knitr)
load("../../datasets/esetUsoskin.RData")
```


```{r cores}
NCORES <- 2
registerDoParallel(NCORES)
register(DoparParam())
```

```{r design}
cellType= droplevels(pData(eset)[,"Level 3"])
batch = pData(eset)[,"Picking sessions"]
counts = exprs(eset)
keep = rowSums(counts>0)>9
counts=counts[keep,]
```


# Methods
## RNA-seq methods
### edgeR
```{r edgeR}

  d <- DGEList(counts)
  d <- suppressWarnings(edgeR::calcNormFactors(d))
  design <- model.matrix(~cellType+batch)
  d <- estimateDisp(d, design)
  fit <- glmFit(d,design)
L <- matrix(0,nrow=ncol(fit$coefficients),ncol=11)
rownames(L) <- colnames(fit$coefficients)
colnames(L) <- c("NF1","NF2","NF3","NF4","NF5","NP1","NP2","NP3","PEP1","PEP2","TH")
L[2:11,1] <- -1/10 #NF1 vs. others
L[2:11,2] <- c(1,rep(-1/10,9)) #NF2 vs. others
L[2:11,3] <- c(-1/10,1,rep(-1/10,8)) #NF3 vs. others
L[2:11,4] <- c(rep(-1/10,2),1,rep(-1/10,7)) #NF4 vs. others
L[2:11,5] <- c(rep(-1/10,3),1,rep(-1/10,6)) #NF5 vs. others
L[2:11,6] <- c(rep(-1/10,4),1,rep(-1/10,5)) #NP1 vs. others
L[2:11,7] <- c(rep(-1/10,5),1,rep(-1/10,4)) #NP2 vs. others
L[2:11,8] <- c(rep(-1/10,6),1,rep(-1/10,3)) #NP3 vs. others
L[2:11,9] <- c(rep(-1/10,7),1,rep(-1/10,2)) #PEP1 vs. others
L[2:11,10] <- c(rep(-1/10,8),1,rep(-1/10,1)) #PEP2 vs. others
L[2:11,11] <- c(rep(-1/10,9),1) #TH vs. others
lrtListEdgeR=list()
for(i in 1:ncol(L)) lrtListEdgeR[[i]] <- zinbwave::glmWeightedF(fit,contrast=L[,i], ZI=FALSE)
padjListEdgeR=lapply(lrtListEdgeR, function(x) p.adjust(x$table$PValue,"BH"))
deGenesEdgeR=unlist(lapply(padjListEdgeR,function(x) sum(x<.05)))
deGenesEdgeR
```

### DESeq2
```{r DESeq2}

  colData <- data.frame(cellType = cellType, batch=batch)
  dse <- DESeqDataSetFromMatrix(countData = counts, colData = colData, design = ~ cellType+batch)
  dse <- estimateSizeFactors(dse, type="poscounts")
  dse <- estimateDispersions(dse, minmu=1e-3)
  dse <- nbinomWaldTest(dse, modelMatrixType="standard", betaPrior=TRUE, minmu=1e-3)
  resultsNames(dse) #for building contrasts, see ?results
L=matrix(0,nrow=length(resultsNames(dse)),ncol=11)
rownames(L)=resultsNames(dse)
L[2:11,1] <- -1/10 #NF1 vs. others
L[2:11,2] <- c(1,rep(-1/10,9)) #NF2 vs. others
L[2:11,3] <- c(-1/10,1,rep(-1/10,8)) #NF3 vs. others
L[2:11,4] <- c(rep(-1/10,2),1,rep(-1/10,7)) #NF4 vs. others
L[2:11,5] <- c(rep(-1/10,3),1,rep(-1/10,6)) #NF5 vs. others
L[2:11,6] <- c(rep(-1/10,4),1,rep(-1/10,5)) #NP1 vs. others
L[2:11,7] <- c(rep(-1/10,5),1,rep(-1/10,4)) #NP2 vs. others
L[2:11,8] <- c(rep(-1/10,6),1,rep(-1/10,3)) #NP3 vs. others
L[2:11,9] <- c(rep(-1/10,7),1,rep(-1/10,2)) #PEP1 vs. others
L[2:11,10] <- c(rep(-1/10,8),1,rep(-1/10,1)) #PEP2 vs. others
L[2:11,11] <- c(rep(-1/10,9),1) #TH vs. others
resList=list()
for(i in 1:ncol(L)) resList[[i]] = results(dse,contrast=L[,i])
lapply(resList, function(x) sum(x$padj<=0.05, na.rm=TRUE))
```

### limma-voom
```{r limma}
  batchLimma=factor(gsub(batch,pattern="-",replacement="")) #valid names
	design <- model.matrix(~ cellType+batchLimma)
	nf <- suppressWarnings(edgeR::calcNormFactors(counts))
	y <- voom(counts, design, plot = FALSE, lib.size = colSums(counts) * nf)
	fit <- lmFit(y, design)
  contrast.matrix = makeContrasts(
    NF1 = -0.1*cellTypeNF2 - 0.1*cellTypeNF3 - 0.1*cellTypeNF4 - 0.1*cellTypeNF5 - 0.1*cellTypeNP1 - 0.1*cellTypeNP2 - 0.1*cellTypeNP3 - 0.1*cellTypePEP1 - 0.1*cellTypePEP2 - 0.1*cellTypeTH,
    NF2 = cellTypeNF2 - 0.1*cellTypeNF3 - 0.1*cellTypeNF4 - 0.1*cellTypeNF5 - 0.1*cellTypeNP1 - 0.1*cellTypeNP2 - 0.1*cellTypeNP3 - 0.1*cellTypePEP1 - 0.1*cellTypePEP2 - 0.1*cellTypeTH,
    NF3 = -0.1*cellTypeNF2 + cellTypeNF3 - 0.1*cellTypeNF4 - 0.1*cellTypeNF5 - 0.1*cellTypeNP1 - 0.1*cellTypeNP2 - 0.1*cellTypeNP3 - 0.1*cellTypePEP1 - 0.1*cellTypePEP2 - 0.1*cellTypeTH,
    NF4 = -0.1*cellTypeNF2 - 0.1*cellTypeNF3 + cellTypeNF4 - 0.1*cellTypeNF5 - 0.1*cellTypeNP1 - 0.1*cellTypeNP2 - 0.1*cellTypeNP3 - 0.1*cellTypePEP1 - 0.1*cellTypePEP2 - 0.1*cellTypeTH,
    NF5 = -0.1*cellTypeNF2 - 0.1*cellTypeNF3 - 0.1*cellTypeNF4 + cellTypeNF5 - 0.1*cellTypeNP1 - 0.1*cellTypeNP2 - 0.1*cellTypeNP3 - 0.1*cellTypePEP1 - 0.1*cellTypePEP2 - 0.1*cellTypeTH,
    NP1 = -0.1*cellTypeNF2 - 0.1*cellTypeNF3 - 0.1*cellTypeNF4 - 0.1*cellTypeNF5 + cellTypeNP1 - 0.1*cellTypeNP2 - 0.1*cellTypeNP3 - 0.1*cellTypePEP1 - 0.1*cellTypePEP2 - 0.1*cellTypeTH,
    NP2 = -0.1*cellTypeNF2 - 0.1*cellTypeNF3 - 0.1*cellTypeNF4 - 0.1*cellTypeNF5 - 0.1*cellTypeNP1 + cellTypeNP2 - 0.1*cellTypeNP3 - 0.1*cellTypePEP1 - 0.1*cellTypePEP2 - 0.1*cellTypeTH,
    NP3 = -0.1*cellTypeNF2 - 0.1*cellTypeNF3 - 0.1*cellTypeNF4 - 0.1*cellTypeNF5 - 0.1*cellTypeNP1 - 0.1*cellTypeNP2 + cellTypeNP3 - 0.1*cellTypePEP1 - 0.1*cellTypePEP2 - 0.1*cellTypeTH,
    PEP1 = -0.1*cellTypeNF2 - 0.1*cellTypeNF3 - 0.1*cellTypeNF4 - 0.1*cellTypeNF5 - 0.1*cellTypeNP1 - 0.1*cellTypeNP2 - 0.1*cellTypeNP3 + cellTypePEP1 - 0.1*cellTypePEP2 - 0.1*cellTypeTH,
    PEP2 = -0.1*cellTypeNF2 - 0.1*cellTypeNF3 - 0.1*cellTypeNF4 - 0.1*cellTypeNF5 - 0.1*cellTypeNP1 - 0.1*cellTypeNP2 - 0.1*cellTypeNP3 - 0.1*cellTypePEP1 + cellTypePEP2 - 0.1*cellTypeTH,
    TH = -0.1*cellTypeNF2 - 0.1*cellTypeNF3 - 0.1*cellTypeNF4 - 0.1*cellTypeNF5 - 0.1*cellTypeNP1 - 0.1*cellTypeNP2 - 0.1*cellTypeNP3 - 0.1*cellTypePEP1 - 0.1*cellTypePEP2 + cellTypeTH,
  levels=design
)
  fit2 <- contrasts.fit(fit, contrast.matrix)
	fit2 <- eBayes(fit2)
  ttList = list()
  for(i in 1:11) ttList[[i]]  <- topTable(fit2, coef = i, n = nrow(counts), sort.by = "none")
  deGenesLimma = lapply(ttList, function(x) sum(x$adj.P.Val<=0.05))
  deGenesLimma
```

## scRNA-seq methods

### MAST

```{r MAST}
counts = exprs(eset)
tpm <- counts*1e6/colSums(counts)
tpm <- log2(tpm+1)
  sca <- FromMatrix(tpm,  cData=data.frame(cellType=cellType, batch=batch))
  #sca <- FromMatrix(counts,  cData=data.frame(group=group))

  # Adaptive thresholding from MAST vignette
  freq_expressed <- 0.1
  thres <- thresholdSCRNACountMatrix(assay(sca), nbins = 200, min_per_bin = 50, conditions = cellType)
  #par(mfrow=c(5,4))
  #plot(thres)
  assays(sca) <- list(thresh=thres$counts_threshold, tpm=assay(sca))
  expressed_genes <- freq(sca) > freq_expressed
  sca <- sca[expressed_genes,]

  ngeneson <- apply(exprs(eset),2,function(x) mean(x>0))
  CD <- colData(sca)
  CD$ngeneson <- ngeneson
  CD$cngeneson <- CD$ngeneson-mean(ngeneson)
  colData(sca) <- CD
  ## differential expression
  fit <- zlm(~ cellType + batch + cngeneson, sca = sca)
  ### contrasts
  L <- matrix(0,nrow=ncol(fit@coefC),ncol=11)
  rownames(L) <- colnames(fit@coefC)
  colnames(L) <- c("NF1","NF2","NF3","NF4","NF5","NP1","NP2","NP3","PEP1","PEP2","TH")
  L[2:11,1] <- -1/10 #NF1 vs. others
  L[2:11,2] <- c(1,rep(-1/10,9)) #NF2 vs. others
  L[2:11,3] <- c(-1/10,1,rep(-1/10,8)) #NF3 vs. others
  L[2:11,4] <- c(rep(-1/10,2),1,rep(-1/10,7)) #NF4 vs. others
  L[2:11,5] <- c(rep(-1/10,3),1,rep(-1/10,6)) #NF5 vs. others
  L[2:11,6] <- c(rep(-1/10,4),1,rep(-1/10,5)) #NP1 vs. others
  L[2:11,7] <- c(rep(-1/10,5),1,rep(-1/10,4)) #NP2 vs. others
  L[2:11,8] <- c(rep(-1/10,6),1,rep(-1/10,3)) #NP3 vs. others
  L[2:11,9] <- c(rep(-1/10,7),1,rep(-1/10,2)) #PEP1 vs. others
  L[2:11,10] <- c(rep(-1/10,8),1,rep(-1/10,1)) #PEP2 vs. others
  L[2:11,11] <- c(rep(-1/10,9),1) #TH vs. others
  lrFitList = list()
  for(i in 1:ncol(L)) lrFitList[[i]] <- lrTest(fit, L[,i,drop=FALSE])
  pval = lapply(lrFitList, function(x) x[, 'hurdle', 'Pr(>Chisq)'])
  padj = lapply(pval, function(x) p.adjust(x,"BH"))
  unlist(lapply(padj, function(x) sum(x<=0.05)))

```

### zinbwave

```{r}
core <- SummarizedExperiment(counts,
                             colData = data.frame(cellType = cellType, batch=batch))
zinb_c <- zinbFit(core, X = '~ cellType + batch', commondispersion = TRUE, epsilon=1e12)
weights = computeObservationalWeights(zinb_c, counts)
```

### zinbwave-edgeR
```{r zinbwaveedger}

  d <- DGEList(counts)
  d <- suppressWarnings(edgeR::calcNormFactors(d))
  design <- model.matrix(~cellType+batch)
  d$weights = weights
  d <- estimateDisp(d, design)
  fit <- glmFit(d,design)
L <- matrix(0,nrow=ncol(fit$coefficients),ncol=11)
rownames(L) <- colnames(fit$coefficients)
colnames(L) <- c("NF1","NF2","NF3","NF4","NF5","NP1","NP2","NP3","PEP1","PEP2","TH")
L[2:11,1] <- -1/10 #NF1 vs. others
L[2:11,2] <- c(1,rep(-1/10,9)) #NF2 vs. others
L[2:11,3] <- c(-1/10,1,rep(-1/10,8)) #NF3 vs. others
L[2:11,4] <- c(rep(-1/10,2),1,rep(-1/10,7)) #NF4 vs. others
L[2:11,5] <- c(rep(-1/10,3),1,rep(-1/10,6)) #NF5 vs. others
L[2:11,6] <- c(rep(-1/10,4),1,rep(-1/10,5)) #NP1 vs. others
L[2:11,7] <- c(rep(-1/10,5),1,rep(-1/10,4)) #NP2 vs. others
L[2:11,8] <- c(rep(-1/10,6),1,rep(-1/10,3)) #NP3 vs. others
L[2:11,9] <- c(rep(-1/10,7),1,rep(-1/10,2)) #PEP1 vs. others
L[2:11,10] <- c(rep(-1/10,8),1,rep(-1/10,1)) #PEP2 vs. others
L[2:11,11] <- c(rep(-1/10,9),1) #TH vs. others
lrtListZinbwaveEdger=list()
for(i in 1:ncol(L)) lrtListZinbwaveEdger[[i]] <- zinbwave::glmWeightedF(fit,contrast=L[,i])
padjListZinbEdgeR=lapply(lrtListZinbwaveEdger, function(x) p.adjust(x$table$PValue,"BH"))
deGenesZinbEdgeR=unlist(lapply(padjListZinbEdgeR,function(x) sum(x<=.05)))
deGenesZinbEdgeR

```

### zinbwave-DESeq2
```{r zinbwavedeseq2}

  colData <- data.frame(cellType = cellType, batch=batch)
  dse <- DESeqDataSetFromMatrix(countData = counts, colData = colData, design = ~ cellType+batch)
  assays(dse)[["weights"]]=weights
  dse <- DESeq2::estimateSizeFactors(dse, type="poscounts")
  dse <- estimateDispersions(dse,minmu=1e-3)
  dse <- nbinomWaldTest(dse, modelMatrixType="standard", betaPrior=TRUE, useT=TRUE, df=rowSums(weights)-13, minmu=1e-3)
resultsNames(dse) #for building contrasts, see ?results
L=matrix(0,nrow=length(resultsNames(dse)),ncol=11)
rownames(L)=resultsNames(dse)
L[2:11,1] <- -1/10 #NF1 vs. others
L[2:11,2] <- c(1,rep(-1/10,9)) #NF2 vs. others
L[2:11,3] <- c(-1/10,1,rep(-1/10,8)) #NF3 vs. others
L[2:11,4] <- c(rep(-1/10,2),1,rep(-1/10,7)) #NF4 vs. others
L[2:11,5] <- c(rep(-1/10,3),1,rep(-1/10,6)) #NF5 vs. others
L[2:11,6] <- c(rep(-1/10,4),1,rep(-1/10,5)) #NP1 vs. others
L[2:11,7] <- c(rep(-1/10,5),1,rep(-1/10,4)) #NP2 vs. others
L[2:11,8] <- c(rep(-1/10,6),1,rep(-1/10,3)) #NP3 vs. others
L[2:11,9] <- c(rep(-1/10,7),1,rep(-1/10,2)) #PEP1 vs. others
L[2:11,10] <- c(rep(-1/10,8),1,rep(-1/10,1)) #PEP2 vs. others
L[2:11,11] <- c(rep(-1/10,9),1) #TH vs. others
resList=list()
for(i in 1:ncol(L)) resList[[i]] = results(dse,contrast=L[,i])
lapply(resList, function(x) sum(x$padj<=0.05, na.rm=TRUE))
```

### zingeR_edgeR

```{r}
d <- DGEList(counts)
d <- suppressWarnings(edgeR::calcNormFactors(d))
design <- model.matrix(~ cellType+batch)
weights <- zingeR::zeroWeightsLS(counts = d$counts, design = design, maxit = 500, normalization = "TMM", verbose = TRUE)
d$weights <- weights
d <- estimateDisp(d, design)
fit <- glmFit(d,design)
L <- matrix(0,nrow=ncol(fit$coefficients),ncol=11)
rownames(L) <- colnames(fit$coefficients)
colnames(L) <- c("NF1","NF2","NF3","NF4","NF5","NP1","NP2","NP3","PEP1","PEP2","TH")
L[2:11,1] <- -1/10 #NF1 vs. others
L[2:11,2] <- c(1,rep(-1/10,9)) #NF2 vs. others
L[2:11,3] <- c(-1/10,1,rep(-1/10,8)) #NF3 vs. others
L[2:11,4] <- c(rep(-1/10,2),1,rep(-1/10,7)) #NF4 vs. others
L[2:11,5] <- c(rep(-1/10,3),1,rep(-1/10,6)) #NF5 vs. others
L[2:11,6] <- c(rep(-1/10,4),1,rep(-1/10,5)) #NP1 vs. others
L[2:11,7] <- c(rep(-1/10,5),1,rep(-1/10,4)) #NP2 vs. others
L[2:11,8] <- c(rep(-1/10,6),1,rep(-1/10,3)) #NP3 vs. others
L[2:11,9] <- c(rep(-1/10,7),1,rep(-1/10,2)) #PEP1 vs. others
L[2:11,10] <- c(rep(-1/10,8),1,rep(-1/10,1)) #PEP2 vs. others
L[2:11,11] <- c(rep(-1/10,9),1) #TH vs. others
lrtListZingerEdgeR=list()
for(i in 1:ncol(L)) lrtListZingerEdgeR[[i]] <- zinbwave::glmWeightedF(fit,contrast=L[,i], ZI=TRUE, independentFiltering=TRUE)
padjListZingerEdgeR=lapply(lrtListZingerEdgeR, function(x) p.adjust(x$table$PValue,"BH"))
deGenesZingerEdgeR=unlist(lapply(padjListZingerEdgeR,function(x) sum(x<.05)))

```

### zingeR_DESeq2

```{r}
  colData <- data.frame(cellType = cellType, batch=batch)
  design <- model.matrix(~ cellType + batch)
  dse <- DESeqDataSetFromMatrix(countData = counts, colData = colData, design = ~cellType+batch)
  weights <- zingeR::zeroWeightsLS(counts = counts, design = design, maxit = 500, normalization = "DESeq2_poscounts", colData = colData, designFormula = ~cellType + batch, verbose = TRUE)
  assays(dse)[["weights"]] <- weights
  dse <- DESeq2::estimateSizeFactors(dse, type="poscounts")
  dse <- estimateDispersions(dse)
  dse <- nbinomWaldTest(dse, modelMatrixType="standard", betaPrior=TRUE, useT=TRUE, df=rowSums(weights)-13, minmu=1e-3)
resultsNames(dse) #for building contrasts, see ?results
L=matrix(0,nrow=length(resultsNames(dse)),ncol=11)
rownames(L)=resultsNames(dse)
L[2:11,1] <- -1/10 #NF1 vs. others
L[2:11,2] <- c(1,rep(-1/10,9)) #NF2 vs. others
L[2:11,3] <- c(-1/10,1,rep(-1/10,8)) #NF3 vs. others
L[2:11,4] <- c(rep(-1/10,2),1,rep(-1/10,7)) #NF4 vs. others
L[2:11,5] <- c(rep(-1/10,3),1,rep(-1/10,6)) #NF5 vs. others
L[2:11,6] <- c(rep(-1/10,4),1,rep(-1/10,5)) #NP1 vs. others
L[2:11,7] <- c(rep(-1/10,5),1,rep(-1/10,4)) #NP2 vs. others
L[2:11,8] <- c(rep(-1/10,6),1,rep(-1/10,3)) #NP3 vs. others
L[2:11,9] <- c(rep(-1/10,7),1,rep(-1/10,2)) #PEP1 vs. others
L[2:11,10] <- c(rep(-1/10,8),1,rep(-1/10,1)) #PEP2 vs. others
L[2:11,11] <- c(rep(-1/10,9),1) #TH vs. others
resList=list()
for(i in 1:ncol(L)) resList[[i]] = results(dse,contrast=L[,i])
lapply(resList, function(x) sum(x$padj<=0.05, na.rm=TRUE))
```

## plot results

```{r resultsData}
results = data.frame(
  nrGenes = c(46 , 106 , 3 , 37 , 105 , 415 , 2 , 3 , 54 , 1 , 381, #scde
            396 , 272 , 529 , 394 , 409 , 1355 , 590 , 241 , 424 , 370 , 1425, #MAST
            6944 , 5128 , 5154 , 801 , 4021 , 2422 , 6724 , 202 , 9388 , 153 , 5735, #limma
            640 , 711 , 186 , 289 , 589 , 1356 , 530 , 316 , 1691 , 75 , 1652, #DESeq2
            522 , 1011 , 348 , 419 , 1216 , 1915 , 762 , 376 , 2203 , 142 , 3087, #ZINB deseq2
            860 , 1375 , 401 , 654 , 1149 , 1942 , 933 , 380 , 1909 , 403 , 2458, #egdeR
            495 , 772 , 241 , 508 , 959 , 1802 , 967 , 464 , 2376 , 193 , 2919), #zinb edger
    contrast=rep(c("NF1","NF2","NF3","NF4","NF5","NP1","NP2","NP3","PEP1","PEP2","TH"),7),
    method=rep(c("SCDE","MAST","limma-voom","DESeq2","ZINB-WaVE_DESeq2", "edgeR","ZINB-WaVE_edgeR"),each=11))
results$method=factor(results$method,levels=c("SCDE","MAST","DESeq2","ZINB-WaVE_DESeq2","edgeR","ZINB-WaVE_edgeR","limma-voom"))

# library(lattice) ; library(grid)
# barchart(log10(nrGenes)~contrast,groups=method,data=results, col=c("grey","darkturquoise","brown","darkseagreen","red","salmon","blue"), page = function(n)
#            grid.text(c("SCDE","MAST", "DESeq2", "ZINB-DESeq2", "edgeR", "ZINB-edgeR", "limma-voom"),
#                      x = seq(0.06,0.9,length=7), y = rep(.9,2),
#                      default.units = "npc"
#                      ,just = c("left", "top")))

library(ggplot2)
# Grouped
ggplot(results, aes(fill=method, y=log(nrGenes), x=contrast)) + geom_bar(position="dodge", stat="identity") +  scale_fill_manual(values = c("grey","darkturquoise","brown","darkseagreen","red","salmon","blue") )+ theme_bw()

# facets
ggplot(results, aes(fill=method, y=log10(nrGenes), x=1)) + geom_bar(position="dodge", stat="identity") + facet_wrap(~contrast) +  scale_fill_manual(values = c("grey","darkturquoise","brown","darkseagreen","red","salmon","blue") ) + theme_bw() + theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank()) + ylab("log10(number of significant genes)")



### including zingeR
resultsZingeR = data.frame(
  nrGenes = c(46 , 106 , 3 , 37 , 105 , 415 , 2 , 3 , 54 , 1 , 381, #scde
            396 , 272 , 529 , 394 , 409 , 1355 , 590 , 241 , 424 , 370 , 1425, #MAST
            6944 , 5128 , 5154 , 801 , 4021 , 2422 , 6724 , 202 , 9388 , 153 , 5735, #limma
            640 , 711 , 186 , 289 , 589 , 1356 , 530 , 316 , 1691 , 75 , 1652, #DESeq2
            522 , 1011 , 348 , 419 , 1216 , 1915 , 762 , 376 , 2203 , 142 , 3087, #ZINB deseq2
            823,  980,  289,  461,  907, 2044,  777,  574, 1933,  145, 2404, #zingeR DESeq2
            860 , 1375 , 401 , 654 , 1149 , 1942 , 933 , 380 , 1909 , 403 , 2458, #egdeR
            495 , 772 , 241 , 508 , 959 , 1802 , 967 , 464 , 2376 , 193 , 2919, #zinb edgeR
            981, 1740,  531,  797, 1542, 2237, 1092,  547, 2053,  492, 2769), #zingeR edger
    contrast=rep(c("NF1","NF2","NF3","NF4","NF5","NP1","NP2","NP3","PEP1","PEP2","TH"),9),
    method=rep(c("SCDE","MAST","limma-voom","DESeq2","ZINB-WaVE_DESeq2","zingeR_DESeq2", "edgeR","ZINB-WaVE_edgeR","zingeR_edgeR"),each=11))
resultsZingeR$method=factor(resultsZingeR$method,levels=c("SCDE","MAST","DESeq2","ZINB-WaVE_DESeq2","zingeR_DESeq2","edgeR","ZINB-WaVE_edgeR","zingeR_edgeR","limma-voom"))


# facets
ggplot(resultsZingeR, aes(fill=method, y=log10(nrGenes), x=1)) + geom_bar(position="dodge", stat="identity") + facet_wrap(~contrast) +  scale_fill_manual(values = c("grey","darkturquoise","brown","darkseagreen","darkolivegreen4","red","salmon","hotpink1","blue") ) + theme_bw() + theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank()) + ylab("log10(number of significant genes)")

```



### compare zingeR-edgeR with edgeR DE genes

```{r}
mains =  c("NF1", "NF2", "NF3", "NF4", "NF5", "NP1", "NP2", "NP3", "PEP1", "PEP2", "TH")
plotVolcanoEdgeRZingeR <- function(comparison){
  plot(x=lrtListZingerEdgeR[[comparison]]$table$logFC, y=-log10(lrtListZingerEdgeR[[comparison]]$table$PValue), xlab="log2 fold change", ylab="- log10(p-value)", main=mains[comparison])
  padjZingeR = p.adjust(lrtListZingerEdgeR[[comparison]]$table$PValue, "BH")
  padjEdgeR = p.adjust(lrtListEdgeR[[comparison]]$table$PValue, "BH")
  colsBoth = (padjZingeR<=0.05 & padjEdgeR<=0.05)
  colsOnlyZinger=(padjZingeR<=0.05 & padjEdgeR>0.05)
  points(x=lrtListZingerEdgeR[[comparison]]$table$logFC[colsBoth], y=-log10(lrtListZingerEdgeR[[comparison]]$table$PValue[colsBoth]), col="blue", pch=16)
  points(x=lrtListZingerEdgeR[[comparison]]$table$logFC[colsOnlyZinger], y=-log10(lrtListZingerEdgeR[[comparison]]$table$PValue[colsOnlyZinger]), col="red", pch=16)
  legend("topleft", c("Significant in edgeR and zingeR_edgeR", "Only significant in zingeR_edgeR", "Non-significant"), pch=c(16,16,1), col=c("blue", "red", "black"), bty="n", cex=1/2)
}

png("~/Dropbox/phdKoen/singleCell/zinbwaveZinger/plots2/volcanoUsoskinZingeR.png", width=9,height=7, units="in", res=300)
library(rafalib) ; mypar()
par(mfrow=c(4,3))
sapply(1:11, plotVolcanoEdgeRZingeR)
dev.off()

plotVolcanoEdgeRZinbwave <- function(comparison){
    plot(x=lrtListZinbwaveEdger[[comparison]]$table$logFC, y=-log10(lrtListZinbwaveEdger[[comparison]]$table$PValue), xlab="log2 fold change", ylab="- log10(p-value)", main=mains[comparison])
    padjZinbwave = p.adjust(lrtListZinbwaveEdger[[comparison]]$table$PValue, "BH")
    padjEdgeR = p.adjust(lrtListEdgeR[[comparison]]$table$PValue, "BH")
    colsBoth = (padjZinbwave<=0.05 & padjEdgeR<=0.05)
    colsOnlyZinger=(padjZinbwave<=0.05 & padjEdgeR>0.05)
    points(x=lrtListZinbwaveEdger[[comparison]]$table$logFC[colsBoth], y=-log10(lrtListZinbwaveEdger[[comparison]]$table$PValue[colsBoth]), col="blue", pch=16)
    points(x=lrtListZinbwaveEdger[[comparison]]$table$logFC[colsOnlyZinger], y=-log10(lrtListZinbwaveEdger[[comparison]]$table$PValue[colsOnlyZinger]), col="red", pch=16)
    legend("topleft", c("Significant in edgeR and ZINBWaVE_edgeR", "Only significant in ZINBWaVE_edgeR", "Non-significant"), pch=c(16,16,1), col=c("blue", "red", "black"), bty="n", cex=1/2)
}

png("~/Dropbox/phdKoen/singleCell/zinbwaveZinger/plots2/volcanoUsoskinZinbwave.png", width=9,height=7, units="in", res=300)
library(rafalib) ; mypar()
par(mfrow=c(4,3))
sapply(1:11, plotVolcanoEdgeRZinbwave)
dev.off()







  plot(x=lrtListZinbwaveEdger[[comparison]]$table$logFC, y=-log10(lrtListZinbwaveEdger[[comparison]]$table$PValue), xlab="log2 fold change", ylab="- log10(p-value)", main=mains[comparison])
  padjZingeR = p.adjust(lrtListZinbwaveEdger[[comparison]]$table$PValue, "BH")
  padjEdgeR = p.adjust(lrtListEdgeR[[comparison]]$table$PValue, "BH")
  colsBoth = (padjZingeR<=0.05 & padjEdgeR<=0.05)
  colsOnlyZinger=(padjZingeR<=0.05 & padjEdgeR>0.05)
  points(x=lrtListZinbwaveEdger[[comparison]]$table$logFC[colsBoth], y=-log10(lrtListZinbwaveEdger[[comparison]]$table$PValue[colsBoth]), col="blue", pch=16)
  points(x=lrtListZinbwaveEdger[[comparison]]$table$logFC[colsOnlyZinger], y=-log10(lrtListZinbwaveEdger[[comparison]]$table$PValue[colsOnlyZinger]), col="red", pch=16)
  legend("topleft", c("Significant in edgeR and ZINBWaVE_edgeR", "Only significant in ZINBWaVE_edgeR", "Non-significant"), pch=c(16,16,1), col=c("blue", "red", "black"), bty="n", cex=1/2)

```













### Check effect of epsilon for ZINB-WaVE edgeR results
```{r}
core <- SummarizedExperiment(counts,
                             colData = data.frame(cellType = cellType, batch=batch))
eps = c(10^-2, 1, 1e2, 1e4, 1e6, 1e8, 1e10, 1e12, 1e14)
epsResultsDE = list()
for(i in 1:length(eps)){
zinbEps <- zinbFit(core, X = '~ cellType + batch', commondispersion = TRUE, epsilon=eps[i])
wEps = computeObservationalWeights(zinbEps, counts)
wEps[wEps<1e-6]=1e-6

  d <- DGEList(counts)
  d <- suppressWarnings(edgeR::calcNormFactors(d))
  design <- model.matrix(~cellType+batch)
  d$weights = wEps
  d <- estimateDisp(d, design)
  fit <- glmFit(d,design)
L <- matrix(0,nrow=ncol(fit$coefficients),ncol=11)
rownames(L) <- colnames(fit$coefficients)
colnames(L) <- c("NF1","NF2","NF3","NF4","NF5","NP1","NP2","NP3","PEP1","PEP2","TH")
L[2:11,1] <- -1/10 #NF1 vs. others
L[2:11,2] <- c(1,rep(-1/10,9)) #NF2 vs. others
L[2:11,3] <- c(-1/10,1,rep(-1/10,8)) #NF3 vs. others
L[2:11,4] <- c(rep(-1/10,2),1,rep(-1/10,7)) #NF4 vs. others
L[2:11,5] <- c(rep(-1/10,3),1,rep(-1/10,6)) #NF5 vs. others
L[2:11,6] <- c(rep(-1/10,4),1,rep(-1/10,5)) #NP1 vs. others
L[2:11,7] <- c(rep(-1/10,5),1,rep(-1/10,4)) #NP2 vs. others
L[2:11,8] <- c(rep(-1/10,6),1,rep(-1/10,3)) #NP3 vs. others
L[2:11,9] <- c(rep(-1/10,7),1,rep(-1/10,2)) #PEP1 vs. others
L[2:11,10] <- c(rep(-1/10,8),1,rep(-1/10,1)) #PEP2 vs. others
L[2:11,11] <- c(rep(-1/10,9),1) #TH vs. others
lrtList=list()
for(i in 1:ncol(L)) lrtList[[i]] <- zinbwave::glmWeightedF(fit,contrast=L[,i])
padjListZinbEdgeR=lapply(lrtList, function(x) p.adjust(x$table$PValue,"BH"))
deGenesZinbEdgeR=unlist(lapply(padjListZinbEdgeR,function(x) sum(x<=.05)))
epsResultsDE[[i]] = deGenesZinbEdgeR
}

```
