---
title: "Impact of the ridge penalty on the performance (TPR and FDP)"
author: "Fanny Perraudeau & Koen Van den Berge"
date: "`r Sys.Date()`"
output:
  html_document:
    fig_height: 7
    fig_width: 7
    toc: yes
    code_folding: hide
    toc_float: yes
---

```{r options, echo=FALSE, results="hide",mesasge=FALSE, error=FALSE, include=FALSE, autodep=TRUE}
knitr::opts_chunk$set(fig.align="center", cache=TRUE, error=FALSE, message=FALSE, warning=TRUE)
library(zinbwave)
library(BiocParallel)
library(doParallel)
library(Biobase)
library(scales)
library(iCOBRA) # roc
library(genefilter) #filtered pvalues
library(RColorBrewer)
library(ggplot2)
#library(wesanderson)
#library(wesanderson)
pathToParentFolder="~/Dropbox/phdKoen/singleCell/zinbwavezingerGitHub/zinbwaveZinger/"
source(paste0(pathToParentFolder,"zingeRsimulationFunctions/simulationHelpFunctions_v7_diffInZero.R"))
```

```{r}
NCORES <- 2
registerDoParallel(NCORES)
register(DoparParam())
```

We want to evaluate the impact of the ridge penalty on the performance (TPR and FDP) using the updated simulation framework from Koen. For the updated simulation framework, I used the code in zinbwaveSimulations/islam_sims_fc2/islam_sims_fc2_newSimulations.Rmd and the functions in zingeRsimulations/simulationHelpFunctions_v7_diffInZero.R



```{r data}
data(islamEset, package = "zingeR")
islamHlp=exprs(islamEset)[-c(1:8),] #first 8 are spike-ins.
cellType=pData(islamEset)[,"cellType"]
paramsIslam = getDatasetMoMPositive(counts = islamHlp)
```

```{r sims}
nSamples=80
grp=as.factor(rep(0:1, each = nSamples/2)) #two-group comparison
nTags=10000 #nr of features
set.seed(11)
DEind = sample(1:nTags,floor(nTags*.1),replace=FALSE) #10% DE
fcSim=(2 + rexp(length(DEind), rate = 1/2)) #fold changes
libSizes=sample(colSums(islamHlp),nSamples,replace=TRUE) #library sizes
simDataIslam <- NBsimSingleCell(foldDiff = fcSim, ind = DEind,
                                dataset = islamHlp, nTags = nTags,
                                group = grp,
                                verbose = TRUE, params = paramsIslam,
                                lib.size = libSizes, cpm="AveLogCPM")
simDataIslam$counts[1:5,1:5]
```

```{r}
# BCV plots
dOrig=suppressWarnings(edgeR::calcNormFactors(DGEList(islamHlp)))
dOrig=estimateGLMTagwiseDisp(estimateGLMCommonDisp(dOrig, design=model.matrix(~cellType), interval=c(0,10)),prior.df=0)

d=suppressWarnings(edgeR::calcNormFactors(DGEList(simDataIslam$counts)))
d=estimateGLMTagwiseDisp(estimateGLMCommonDisp(d, design=model.matrix(~grp), interval=c(0,10)),prior.df=0)
```

```{r}
par(mfrow=c(1,2))
plotBCV(dOrig,ylim=c(0,20), xlim=c(0,20))
plotBCV(d,ylim=c(0,20), xlim=c(0,20))
par(mfrow=c(1,1))
```

```{r}
# association of library size with zeros
plot(x=log(colSums(islamHlp)), y=colMeans(islamHlp==0), xlab="Log library size", ylab="Fraction of zeros", xlim=c(8,20))
points(x=log(colSums(simDataIslam$counts)), y=colMeans(simDataIslam$counts==0), col=2)
```

```{r}
# association of aveLogCPM with zeros
plot(x=edgeR::aveLogCPM(islamHlp), y=rowMeans(islamHlp==0), xlab="Average log CPM", ylab="Fraction of zeros", ylim=c(0,1), col=alpha(1,1/2), pch=19, cex=.3)
points(x=edgeR::aveLogCPM(simDataIslam$counts), y=rowMeans(simDataIslam$counts==0),col=alpha(2,1/2),pch=19,cex=.3)
```

# Methods

zinbwave for different epsilon

```{r zinbwaveWeights}
computeZinbwaveWeights <- function(zinb, counts){
  mu <- getMu(zinb)
  pi <- getPi(zinb)
  theta <- getTheta(zinb)
  theta_mat <- matrix(rep(theta, each = ncol(counts)), ncol = nrow(counts))
  nb_part <- dnbinom(t(counts), size = theta_mat, mu = mu)
  zinb_part <- pi * ( t(counts) == 0 ) + (1 - pi) *  nb_part
  zinbwg <- ( (1 - pi) * nb_part ) / zinb_part
  t(zinbwg)
}
```

```{r}
zinbwave_edgeR <- function(counts, group, zinb){
  library(zingeR)
  d=DGEList(counts)
  d=suppressWarnings(calcNormFactors(d))
  design=model.matrix(~group)
  weights <- computeZinbwaveWeights(zinb, d$counts)
  d$weights <- weights
  d=estimateDisp(d, design)
  plotBCV(d)
  fit=glmFit(d,design)
  lrt=glmWeightedF(fit,coef=2, independentFiltering = TRUE)
  cbind(pval = lrt$table$PValue, padj = lrt$table$padjFilter,
        logFC = lrt$table$logFC)
}
```

```{r}
zinbwave_DESeq2 <- function(counts, group, zinb){
  colData=data.frame(group=group)
  design=model.matrix(~group)
  dse=DESeqDataSetFromMatrix(countData=counts, colData=colData, design=~group)
  weights <- computeZinbwaveWeights(zinb, counts)
  #weights[weights<1e-6]=1e-6 #temporary fix.
  assays(dse)[["weights"]]=weights
  dse = DESeq2::estimateSizeFactors(dse, type="poscounts")
  dse = estimateDispersions(dse)
  dse = nbinomWaldTest(dse, betaPrior=TRUE, useT=TRUE, df=rowSums(weights)-2)
  res = results(dse)
  cbind(pval = res$pvalue, padj = res$padj, logFC = res$log2FoldChange)
}
```


```{r core}
core <- SummarizedExperiment(simDataIslam$counts,
                             colData = data.frame(grp = grp))
```

```{r zinbcommondisp}
epsVec = c(10^-2, 1, 1e2, 1e4, 1e6, 1e8, 1e10, 1e12, 1e14)
# zinbList <- lapply(epsVec, function(eps){
#   print(eps)
#   zinbFit(core, X = '~ grp', commondispersion = TRUE, epsilon = eps)
# })
```

# Results

## Biological coefficient of variation (BCV)

```{r, warning=FALSE}
counts = simDataIslam$counts
group = grp

# par(mfrow = c(3, 3))
# resList = lapply(zinbList, function(zinb){
#   zinbwave_edgeR(counts, grp, zinb)
# })
# par(mfrow = c(1, 1))
# save(resList,file="resListEdgerIslam.rda")
#
# resListDESeq2 = lapply(zinbList, function(zinb){
#   zinbwave_DESeq2(counts, grp, zinb)
# })
# save(resListDESeq2,file="resListDESeq2Islam.rda")

load("~/resListEdgerIslam.rda")
load("~/resListDESeq2Islam.rda")

```

## pval histograms
```{r}
res = lapply(resList, as.data.frame)
names(res) = as.character(epsVec)

resDESeq =  lapply(resListDESeq2, as.data.frame)
names(resDESeq) = as.character(epsVec)
```

```{r histogramPVAL}
## edgeR
par(mfrow = c(3, 3))
hist = lapply(1:length(res), function(i){
  hist(res[[i]][,'pval'], main = paste0('epsilon=', names(res)[i]), xlab="p-value", breaks=seq(0,1,0.05))
})
par(mfrow = c(1, 1))

## DESeq2
par(mfrow = c(3, 3))
hist = lapply(1:length(resDESeq), function(i){
  hist(resDESeq[[i]][,'pval'], main = paste0('epsilon=', names(resDESeq)[i]), xlab="p-value", breaks=seq(0,1,0.05))
})
par(mfrow = c(1, 1))
```

## Volcano plots
```{r volcanos}
trueDE = rep(0, nTags)
trueDE[simDataIslam$indDE] = 1

par(mfrow = c(3, 3))
volcanos <- lapply(1:length(res), function(i){
  plot(res[[i]][, 'logFC'], -log10(res[[i]][, 'padj']),
       pch = 20, col = 'gray', cex = .5, xlim = c(-10, 10), ylim = c(0, 15),
       ylab = '-log10(pvalue)', xlab = 'logFC',
       main = paste0('epsilon=',names(res)[i]))
  points(res[[i]][simDataIslam$indDE, 'logFC'], pch=1, col=3, cex=.5, lwd=1,
          -log10(res[[i]][simDataIslam$indDE, 'padj']))
})
par(mfrow = c(1, 1))

par(mfrow = c(3, 3))
volcanos <- lapply(1:length(resDESeq), function(i){
  plot(resDESeq[[i]][, 'logFC'], -log10(resDESeq[[i]][, 'padj']),
       pch = 20, col = 'gray', cex = .5, xlim = c(-10, 10), ylim = c(0, 15),
       ylab = '-log10(pvalue)', xlab = 'logFC',
       main = paste0('epsilon=',names(resDESeq)[i]))
  points(resDESeq[[i]][simDataIslam$indDE, 'logFC'], pch=1, col=3, cex=.5, lwd=1,
          -log10(resDESeq[[i]][simDataIslam$indDE, 'padj']))
})
par(mfrow = c(1, 1))

```

```{r}
lapply(resList, function(x){
  sum(is.na(x[, 'padj']))
})
lapply(resListDESeq2, function(x){
  sum(is.na(x[, 'padj']))
})
```

## nDE, TPR, FDR at pval=0.05
```{r nb}
#edgeR
tab_de = lapply(res, function(y){
  nDE = sum(y$padj <= 0.05, na.rm = TRUE)
  TPR = mean(simDataIslam$indDE %in% which( y$padj <= 0.05))
  FPR = mean(which(y$padj <= 0.05) %in% simDataIslam$indNonDE)
  c(nDE = nDE, TPR = TPR, FPR = FPR)
})
do.call(rbind, tab_de)

#DESeq2
tab_de = lapply(resDESeq, function(y){
  nDE = sum(y$padj <= 0.05, na.rm = TRUE)
  TPR = mean(simDataIslam$indDE %in% which( y$padj <= 0.05))
  FPR = mean(which(y$padj <= 0.05) %in% simDataIslam$indNonDE)
  c(nDE = nDE, TPR = TPR, FPR = FPR)
})
do.call(rbind, tab_de)
```

## TPR vs FDR
```{r plot}
pp = COBRAData(pval = as.data.frame(do.call(cbind, lapply(res, '[[', 1))),
               padj = as.data.frame(do.call(cbind, lapply(res, '[[', 2))),
               truth = data.frame(status = trueDE))
cobraperf <- calculate_performance(pp, binary_truth = "status", thrs = 0.05)
cobraplot <- prepare_data_for_plot(cobraperf)
png("~/Dropbox/phdKoen/singleCell/zinbwaveZinger/plots2/islamEpsEval.png", width=7,height=8, units="in", res=300)
plot_fdrtprcurve(cobraplot, plottype = c("curve", "points"), pointsize = 1,linewidth = .5) + xlab("FDP")
dev.off()

pp = COBRAData(pval = as.data.frame(do.call(cbind, lapply(resDESeq, '[[', 1))),
               padj = as.data.frame(do.call(cbind, lapply(resDESeq, '[[', 2))),
               truth = data.frame(status = trueDE))
cobraperf <- calculate_performance(pp, binary_truth = "status", thrs = 0.05)
cobraplotDESeq <- prepare_data_for_plot(cobraperf)
plot_fdrtprcurve(cobraplot, plottype = c("curve", "points"), pointsize = 1,linewidth = .5) + xlab("FDP")
```

```{r}
plot_fdrtprcurve(cobraplot, plottype = c("curve", "points"), pointsize = .1, xaxisrange = c(0, 0.2),linewidth = .5) + xlab("FDP")

plot_fdrtprcurve(cobraplotDESeq, plottype = c("curve", "points"), pointsize = .1, xaxisrange = c(0, 0.2),linewidth = .5) + xlab("FDP")

```

