---
title: "Case study on Usoskin dataset - FPR on mocks"
author: "Fanny Perraudeau"
date: "`r Sys.Date()`"
output: 
  html_document: 
    fig_height: 7
    fig_width: 7
    toc: yes
    code_folding: hide
    toc_float: yes
---

```{r options, echo=FALSE, results="hide",mesasge=FALSE, error=FALSE, include=FALSE, autodep=TRUE}
knitr::opts_chunk$set(fig.align="center", cache=TRUE, error=FALSE, message=FALSE, warning=TRUE)
library(edgeR)
library(zinbwave)
```

We want to evaluate the impact of the ridge penalty ($\epsilon$) of ZINB-WaVE model on the false positive rate (FPR) for the mock comparisons from the Usoskin dataset.

# Data

Dataset is from the supplementary data accompanying the original paper downloaded at http://linnarssonlab.org/drg/.

```{r data}
path = '../../datasets/usoskin/'
load(paste0(path, 'esetUsoskin.RData'))
eset=eset[rowSums(exprs(eset)>0)>=20,]
exprs(eset) <- apply(exprs(eset),2,function(x) {storage.mode(x) <- 'integer'; x})

file = paste0(path, "subsetMatrixUsoskinFPR_randomCellTypes.txt")
subsets <- read.table(file)
```

# Influence of epsilon

## Compute ZINB-WaVE observational weights

We add the batch (picking session) as a covariate in the ZINB-WaVE model.

```{r dataEloop}
i = 1 
eLoop <- eset[,as.numeric(subsets[i,])]
cat('Removing ', sum(rowSums(exprs(eLoop)) == 0), " genes with only 0's")
eLoop <- eLoop[rowSums(exprs(eLoop)) != 0, ]
condition=factor(rep(c("A","B"),each=45))
pickingSession=factor(rep(rep(c("Cold","rt1","rt2"),each=15),2))
pData(eLoop)$condition=condition
pData(eLoop)$pickingSession=pickingSession
design <- model.matrix(~ condition + pickingSession)
epsVec = 10^seq(0, 14, by = 2)
```

```{r zinbesp,eval=FALSE}
library(BiocParallel)
library(doParallel)
NCORES <- 2
registerDoParallel(NCORES)
register(DoparParam())

zinbList <- lapply(epsVec, function(eps){
  zinbFit(exprs(eLoop), X = design, epsilon = eps)
})
save(zinbList, file = 'zinbList_epsilon.rda')
```

```{r}
load('zinbList_epsilon.rda')
zinbList = zinbList[1:length(epsVec)]
```

```{r pvals}
weightsList <- lapply(zinbList, function(x){
  computeObservationalWeights(x, exprs(eLoop))
})
```

## Compute p-values

We use egdeR with ZINB-WaVE posterior probabilities as observation weights.

```{r}
pvalsList <- lapply(weightsList, function(w){
  d <- DGEList(exprs(eLoop))
  d <- edgeR::calcNormFactors(d)
  d$weights <- w
  d=estimateDisp(d, design)
  fit=glmFit(d,design)
  lrt=glmWeightedF(fit, coef=2, independentFiltering = TRUE)
  lrt$table$PValue
})
```

## Biological coefficient of variation (BCV)
```{r mocksEspilonBCV}
par(mfrow=c(3,4))
myplot <- lapply(1:length(weightsList), function(i){
  d <- DGEList(exprs(eLoop))
  d <- edgeR::calcNormFactors(d)
  w = weightsList[[i]]
  w[is.na(w)] = 1
  d$weights <- w
  d=estimateDisp(d, design)
  plotBCV(d, main = paste0('epsilon=', epsVec[i]), ylim = c(0,6))
})
par(mfrow=c(1,1))
```

## Posterior probabilities
```{r mocksEspilonWeights}
par(mfrow = c(3,4))
hh = lapply(1:length(weightsList), function(i){
   hist(weightsList[[i]][exprs(eLoop)==0], main = paste0('epsilon=', epsVec[i]), ylim = c(0,5e5))
})
par(mfrow = c(1,1))
```

## FPR
```{r fdrEpsUsoskinEps}
fpr = sapply(pvalsList, function(x) mean(x <= 0.05))
print(length(fpr))
plot(log10(epsVec), fpr, main = '', type = 'o', 
     xlab = expression(paste(epsilon, ' (log10)')), ylab = 'PCER',cex.lab = 1.5,cex.axis= 1.5)
abline(h = 0.05, col = 'red')
```

## Histogram of p-values.

We are expecting uniform distribution for the p-values as the 2 groups contain cells that have been randomly sampled from all the cells. Therefore, the true number of DE genes is null. When $\epsilon < 10^6$, p-values are not uniformly distributed.

```{r fdrEpsUsoskinPval}
par(mfrow = c(3,3))
hh = lapply(1:length(weightsList), function(i){
  hist(pvalsList[[i]], main = paste0('epsilon=', epsVec[i]), ylim=c(0,2000),cex.lab = 1.5,cex.axis= 1.5,cex.main=2, xlab='p-values')
})
par(mfrow = c(1,1))
```

## Plot paper

```{r fdrEpsUsoskin,fig.width=10,fig.height=5}
par(mar = c(4.1, 4.25, 3, 1), bty = "l")
layout(cbind(matrix(1,3,3), matrix(2:10,3,3,byrow = TRUE)))
# pcer = f(eps)
fpr = sapply(pvalsList, function(x) mean(x <= 0.05))
print(length(fpr))
plot(log10(epsVec), fpr, main = '', type = 'o',
     xlab = expression(paste(epsilon,' (log10)')), ylab = 'PCER',cex.lab = 1.5,cex.axis= 1.5,cex.main=1.5)
abline(h = 0.05, col = 'red')
mtext("a" ,at=-2, font=2, padj = -1,cex=4/3)

# hist pvalues
hh = lapply(1:length(epsVec), function(i){
  hist(pvalsList[[i]], main = paste0('epsilon=', epsVec[i]), ylim=c(0,2000),cex.lab = 1.5,cex.axis= 1.5,cex.main=1.5, xlab = 'p-values')
  if (i==1) mtext("b" ,at=-1,padj = -1, font=2, cex=4/3)
})
```


## Conclusion

$\epsilon>10^6$ seems to be reasonable.

# sessionInfo()
```{r}
sessionInfo()
```

