---
title: "DE for CD4+ Tcells - clustering on PCA"
author: "Fanny Perraudeau"
date: "`r Sys.Date()`"
output: 
html_document: 
fig_height: 7
fig_width: 7
toc: yes
code_folding: hide
toc_float: yes
---

```{r options, echo=FALSE, results="hide",mesasge=FALSE, error=FALSE, include=FALSE, autodep=TRUE}
knitr::opts_chunk$set(fig.align="center", cache=TRUE, error=FALSE, message=FALSE, warning=TRUE)
library(zinbwave)
library(edgeR)
library(RColorBrewer)
library(ggplot2)
source('../../functionsToZinbwavePackage.R')
```

Using clustering on PCA, we want to compare the different DE methods and perform gene set enrichment analysis.

# Data
```{r loadobject}
load("../data/core.rda")
core = core[,colData(core)$seurat %in% 1:2]
core = core[rowSums(assay(core)) > 0, ]
colData(core)$seurat = factor(colData(core)$seurat)
core
```

# Compute ZINB-WaVE observational weights

```{r,eval=FALSE}
#previous epsilon=1e8 but then looked at mocks
library(doParallel)
library(BiocParallel)
NCORES = 2
registerDoParallel(NCORES)
register(DoparParam())
print(system.time(zinb <- zinbFit(core, X = '~ seurat',
                                  epsilon = 1e12)))
save(zinb, file = 'zinb.rda')
```

```{r weightszinb}
load('zinb.rda')
counts = assay(core)
weights_zinbwave = computeObservationalWeights(zinb, counts)
```

```{r weightszinbZeros}
hist(weights_zinbwave[assay(core) == 0], main = 'zinbwave',
     xlab = 'Weights')
```

```{r}
design = model.matrix(~ colData(core)$seurat)
```

# Methods
## edgeR
```{r fitedger}
fit_edgeR <- function(counts, design, filter = NULL){
  library(edgeR)
  d = DGEList(counts)
  d = suppressWarnings(calcNormFactors(d))
  d = estimateDisp(d, design)
  fit = glmFit(d, design)
  glm = glmLRT(fit)
  tab = glm$table
  tab$padj = p.adjust(tab$PValue, "BH")
  tab$gene = rownames(tab)
  de <- as.data.frame(tab, stringsAsFactors = FALSE)
  de = de[, c('gene', 'PValue', 'padj', 'logFC')]
  colnames(de) = c('gene', 'pval', 'padj', 'logfc')
  de
}
```

```{r edgeR}
edgeR <- fit_edgeR(counts, design)
edgeR$method <- 'edgeR'
```

## zinbwave-weighted edgeR
```{r fitedgeRzi}
fit_edgeR_zi <- function(counts, design, weights,
                         filter = NULL){
  library(edgeR)
  d = DGEList(counts)
  d = suppressWarnings(calcNormFactors(d))
  d$weights <- weights 
  d = estimateDisp(d, design)
  fit = glmFit(d,design)
  glm = glmWeightedF(fit, filter = filter)
  tab = glm$table
  tab$gene = rownames(tab)
  de <- data.frame(tab, stringsAsFactors = FALSE)
  de = de[, c('gene', 'PValue', 'padjFilter', 'logFC')]
  colnames(de) = c('gene', 'pval', 'padj', 'logfc')
  de
}
```

```{r edgeRzi}
nf <- edgeR::calcNormFactors(counts)
baseMean = unname(rowMeans(sweep(counts,2,nf,FUN="*")))
zinbwave_edgeR <- fit_edgeR_zi(counts, design, 
                               weights = weights_zinbwave,
                               filter = baseMean)
zinbwave_edgeR$method <- 'zinbwave_edgeR'
```

## limma-voom
```{r runlimmavoom}
runLimmavoom <- function(counts, design) {
  library(limma)
  library(edgeR)
  dgel <- DGEList(counts)
  dgel <- edgeR::calcNormFactors(dgel)
  v <- voom(dgel,design,plot=FALSE)
  fit <- lmFit(v,design)
  fit <- eBayes(fit)
  tt <- topTable(fit,coef=2,n=nrow(dgel),sort.by="none")
  pvals <- tt$P.Value
  padj <- p.adjust(pvals,method="BH")
  padj[is.na(padj)] <- 1
  data.frame(gene = rownames(tt), pval=pvals, padj=padj, logfc=tt$logFC)
}
```

```{r limmavoom}
voom <- runLimmavoom(counts, design)
voom$method <- 'limmavoom'
```

# Seurat
```{r seurat}
seurat = read.csv('tcellmarkers_seurat.csv')
seurat$method = 'seurat'
seurat = seurat[, c('gene', 'p_val', 'p_val_adj',
                    'avg_logFC', 'method')]
colnames(seurat) = colnames(edgeR)
```

# Results
Let's compare the different methods.

```{r res}
de = rbind(edgeR, seurat, voom, zinbwave_edgeR)
write.csv(de, file = 'deGenes.csv', row.names = FALSE)
```

## GSEA pre ranked
We use gene set enrichment analysis and xCell gene sets.

```{r}
library(xCell) # for db
library(fgsea)
library(GSEABase)

## extract genesets from xcell
genesets <- unlist(geneIds(xCell.data$signatures))
celltypes <- sapply(strsplit(names(genesets), "%"), function(x) x[1])
names(genesets) <- NULL
gs <- tapply(genesets, celltypes, c)
de$logfc[is.na(de$logfc)] = rnorm(n = sum(is.na(de$logfc)), 0, 0.005) 
# MAST creates NAs for log fold change but fgsea does not handle NAs, so I artificially creates small log fold change. rnorm with mean = 0, and not 0 because fgsea does not handle tights.
set.seed(6372)
gsea_res = lapply(unique(de$method), function(x){
  print(x)
  temp = de[de$method == x, ]
  degenes = temp$logfc
  if (x == 'seurat') degenes = - degenes # reverse design matrix in seurat
  names(degenes) = temp$gene
  gsea = fgsea(gs, degenes, nperm = 10000, minSize = 5)
  gsea$method = x
  gsea[order(-abs(gsea$NES)), ]
})
lapply(gsea_res, head)
```

```{r}
gseaDf = as.data.frame(do.call(rbind, gsea_res))
gseaDf = gseaDf[gseaDf$size > 100, ]
gseaDf = gseaDf[, c('method', 'pathway', 'NES')]
#gseaDf$method = factor(gseaDf$method, levels = c('edgeR', 'seurat', 'MAST', 'zinbwave_DESeq2', 'limmavoom', 'zinbwave_edgeR'))
sortedPwy = gseaDf[gseaDf$method == 'zinbwave_edgeR', ]
sortedPwy = sortedPwy[order(sortedPwy$NES), 'pathway']
gseaDf$pathway = factor(gseaDf$pathway, levels = sortedPwy)
```

```{r tenxcaseNESallPCA}
ggplot(gseaDf, aes(method, pathway)) +
  geom_tile(aes(fill = NES)) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_fill_gradient2(low = "blue", high = "red",
                       mid = "white", midpoint = 0,
                       space = "Lab", 
                       name="Normalized\nEnrichment\nScore") +
  ylab('Cell Type') + xlab('Method')
```

```{r tenxcaseNEScd4PCA}
size=20
chosen = c('CD4+ memory T-cells','CD4+ Tem','CD4+ Tcm','CD4+ naive T-cells')
sub = gseaDf[gseaDf$pathway %in% chosen, ]
nes = sub$NES
limit = max(abs(min(nes)), max(nes))
sub$pathway = factor(sub$pathway, levels = chosen)
ggplot(sub, aes(method, pathway)) +
  geom_tile(aes(fill = NES)) + 
  scale_fill_gradient2(low = "blue", high = "red", mid = "white", 
                       midpoint = 0, space = "Lab",
                       name="Normalized\nEnrichment\nScore", 
                       limit = c(-limit, limit)) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  ylab('Cell Type') + xlab('Method') +
  theme(text = element_text(size = size))
```

```{r}
lapply(gsea_res, function(x){
  x[x$pathway %in% chosen, ]
})
```


## Histogram of pvalues
We are expecting uniformity of the pvalue with a pick close to zero corresponding to DE genes.

```{r tenxcaseHistPvalPCA}
par(mfrow= c(2,2))
for (x in unique(de$method)){
  hist(de[de$method == x, 'pval'], main = x, ylim = c(0, 3000),
       xlab = 'pvalue')
}
par(mfrow= c(1,1))
```

## Number of DE genes
```{r}
for (x in unique(de$method)){
  print(x)
  print(sum(de[de$method ==x, 'pval'] < 0.05, na.rm = TRUE))
  print(sum(de[de$method ==x, 'padj'] < 0.05, na.rm = TRUE))
}
```

## Concordance between DE genes
### Venn diagram

```{r}
thr = 0.05
```

DE genes are genes with an adjusted pvalue lower than `r thr`.

```{r venn}
library(dplyr)
ve = de %>% group_by(method, gene) %>%
  summarize(pval = min(pval, na.rm=TRUE)) %>%
  ungroup() %>% as.data.frame()
ve$de = ve$pval < thr
ve = reshape2::acast(ve[, c(1,2,4)], gene ~ method, sum)

aa <- vennCounts(ve)
vennDiagram(aa, main = 'DE gene, adj pvalue < 0.05')
```


## Most DE genes per method in terms of pvalue
```{r}
ranks = lapply(unique(de$method), function(x){
  temp = de[de$method == x, ]
  temp$rank = rank(temp$pval)
  temp
})
de = do.call(rbind, ranks)

ranks = lapply(unique(de$method), function(x){
  temp = de[de$method == x, ]
  temp = temp[temp$rank < 10, ]
  temp[order(temp$rank), ]
})
names(ranks) = unique(de$method)
ranks
```

## Most DE genes per method in terms of fold change
```{r}
ranks = lapply(unique(de$method), function(x){
  temp = de[de$method == x, ]
  temp$rank = rank(-abs(temp$logfc))
  temp
})
de = do.call(rbind, ranks)

ranks = lapply(unique(de$method), function(x){
  temp = de[de$method == x, ]
  temp = temp[temp$rank < 10, ]
  temp[order(temp$rank), ]
})
names(ranks) = unique(de$method)
ranks
```


#sessionInfo
```{r}
sessionInfo()
```
