---
title: "DE for CD4+ Tcells - Seurat DE method"
output:
  html_document:
    theme: united
date: 'Compiled: `r format(Sys.Date(), "%B %d, %Y")`'
---
***

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  cache = TRUE,
  cache.lazy = FALSE,
  tidy = TRUE
)
```

When we load package Seurat, many other packages are loaded. Thus, when we want to load other packages to compare DE methods, the maximal number of DLLs is reached. That's the reason why we perform DE analysis in a separate file for Seurat method.

# Data

Data file `pbmc3k.rda` has been created from running `createDataObject.Rmd`.

```{r}
library(Seurat)
load("../data/pbmc3k.rda")
```

# Find DE genes

We use the default `wilcox` test from Seurat package.

```{r}
tcell.markers <- FindMarkers(object = pbmc,
                             ident.1 = 0,
                             ident.2 = 1,
                             logfc.threshold = 0,
                             genes.use = rownames(pbmc@raw.data),
                             thresh.use = 0,
                             test.use = "wilcox",
                             min.pct = 0, 
                             min.diff.pct = -Inf,
                             print.bar = TRUE,
                             only.pos = FALSE,
                             max.cells.per.ident = Inf, 
                             random.seed = 1,
                             latent.vars = "nUMI",
                             min.cells = 0 )
```

```{r restore}
library(dplyr)
head(tcell.markers, 2)
length(unique(rownames(pbmc@raw.data)))
length(unique(rownames(tcell.markers)))
tcell.markers$gene = rownames(tcell.markers)
topgenes = tcell.markers %>% top_n(20, -p_val)
topgenes
```

Let's look at some markers.
```{r}
FeaturePlot(object = pbmc, features.plot = c("IL32", "S100A4"),
            cols.use = c("green", "blue"))
```

```{r}
FeaturePlot(object = pbmc, features.plot = c("S100A4", "CCR7"),
            cols.use = c("green", "blue"))
```

```{r}
write.csv(tcell.markers, file = 'tcellmarkers_seurat.csv')
```

#sessionInfo
```{r}
sessionInfo()
```
