---
title: "Impact of the ridge penalty on the performance (TPR and FDP)"
author: "Fanny Perraudeau & Koen Van den Berge"
date: "`r Sys.Date()`"
output:
  html_document:
    fig_height: 7
    fig_width: 7
    toc: yes
    code_folding: hide
    toc_float: yes
---

```{r options, echo=FALSE, results="hide",mesasge=FALSE, error=FALSE, include=FALSE, autodep=TRUE}
knitr::opts_chunk$set(fig.align="center", cache=TRUE, error=FALSE, message=FALSE, warning=TRUE)
library(zinbwave)
library(BiocParallel)
library(doParallel)
library(Biobase)
library(scales)
library(iCOBRA) # roc
library(genefilter) #filtered pvalues
library(RColorBrewer)
library(ggplot2)
pathToParentFolder="~/Dropbox/phdKoen/singleCell/zinbwavezingerGitHub/zinbwaveZinger/"
source(paste0(pathToParentFolder,"zingeRsimulationFunctions/simulationHelpFunctions_v7_diffInZero.R"))
```

```{r}
NCORES <- 2
registerDoParallel(NCORES)
register(DoparParam())
```

We want to evaluate the impact of the ridge penalty on the performance (TPR and FDP) using the updated simulation framework from Koen. For the updated simulation framework, I used the code in zinbwaveSimulations/islam_sims_fc2/islam_sims_fc2_newSimulations.Rmd and the functions in zingeRsimulations/simulationHelpFunctions_v7_diffInZero.R



```{r data}
install_cellranger = FALSE
if (install_cellranger){
  source("http://cf.10xgenomics.com/supp/cell-exp/rkit-install-2.0.0.R")
  library(cellrangerRkit)
  packageVersion("cellrangerRkit")
}
library(cellrangerRkit)
pipestance_path <- "./"
if(!file.exists(paste0(pipestance_path, "/outs"))) {
  download_sample(sample_name="pbmc4k",sample_dir=pipestance_path,
                  host="http://cf.10xgenomics.com/samples/cell-exp/1.3.0/")
}
pbmc <- load_cellranger_matrix(pipestance_path)
analysis_results <- load_cellranger_analysis_results(pipestance_path)
counts = exprs(pbmc) # expression matrix
cl <- analysis_results[["clustering"]][["graphclust"]][,'Cluster']
names(cl) <- analysis_results[["clustering"]][["graphclust"]][,'Barcode']
cl <- cl[colnames(pbmc)]
counts = counts[, cl %in% 2:3]
cellType = cl[cl %in% 2:3]
counts <- counts[rowSums(counts != 0) > 20, ] #expression in at least 1/60 cells
tenXData = counts

paramsTenx = getDatasetMoMPositive(counts = tenXData)
```

```{r sims}
nSamples=ncol(tenXData)
grp=as.factor(rep(0:1, each = nSamples/2)) #two-group comparison
nTags=10000 #nr of features
set.seed(11)
DEind = sample(1:nTags,floor(nTags*.1),replace=FALSE) #10% DE
fcSim=(2 + rexp(length(DEind), rate = 1/2)) #fold changes
libSizes=sample(colSums(tenXData),nSamples,replace=TRUE) #library sizes
simData <- NBsimSingleCell(foldDiff = fcSim, ind = DEind,
                           dataset = tenXData, nTags = nTags,
                           group = grp,
                           verbose = TRUE, params = paramsTenx,
                           lib.size = libSizes, cpm="AveLogCPM", normalizeLambda=TRUE, min.dispersion=1e-3)

```

# Methods

zinbwave for different epsilon

```{r}
zinbwave_edgeR <- function(counts, group, zinb){
  d=DGEList(counts)
  d=suppressWarnings(calcNormFactors(d))
  design=model.matrix(~group)
  weights <- computeObservationalWeights(zinb, d$counts)
  weights[weights==0]=1e-8
  d$weights <- weights
  d=estimateDisp(d, design)
  plotBCV(d)
  fit=glmFit(d,design)
  lrt=glmWeightedF(fit,coef=2, independentFiltering = TRUE)
  cbind(pval = lrt$table$PValue, padj = lrt$table$padjFilter,
        logFC = lrt$table$logFC)
}
```

```{r core}
core <- SummarizedExperiment(simData$counts,
                             colData = data.frame(grp = grp))
```

```{r zinbcommondisp}
epsVec = c(10^-2, 1, 1e2, 1e4, 1e6, 1e8, 1e10, 1e12, 1e14)
zinbList <- lapply(epsVec, function(eps){
  print(eps)
  zinbFit(core, X = '~ grp', commondispersion = TRUE, epsilon = eps)
})
```

# Results

## Biological coefficient of variation (BCV)

```{r, warning=FALSE}
counts = simData$counts
group = grp

par(mfrow = c(3, 3))
resList = lapply(zinbList, function(zinb){
  zinbwave_edgeR(counts, grp, zinb)
})
par(mfrow = c(1, 1))
save(resList,file="resListEdger10x.rda")


load("~/resListEdger10x.rda")

```

## pval histograms
```{r}\
for(i in 1:length(resList)){
  resList[[i]][is.na(resList[[i]][,"pval"]),"pval"] = 1
  resList[[i]][is.na(resList[[i]][,"padj"]),"padj"] = 1
}

res = lapply(resList, as.data.frame)
names(res) = as.character(epsVec)

# resDESeq =  lapply(resListDESeq2, as.data.frame)
# names(resDESeq) = as.character(epsVec)
```

```{r histogramPVAL}
## edgeR
par(mfrow = c(3, 3))
hist = lapply(1:length(res), function(i){
  hist(res[[i]][,'pval'], main = paste0('epsilon=', names(res)[i]), xlab="p-value", breaks=seq(0,1,0.05))
})
par(mfrow = c(1, 1))

# ## DESeq2
# par(mfrow = c(3, 3))
# hist = lapply(1:length(resDESeq), function(i){
#   hist(resDESeq[[i]][,'pval'], main = paste0('epsilon=', names(resDESeq)[i]), xlab="p-value", breaks=seq(0,1,0.05))
# })
# par(mfrow = c(1, 1))
```

## Volcano plots
```{r volcanos}
trueDE = rep(0, nTags)
trueDE[simDataIslam$indDE] = 1

par(mfrow = c(3, 3))
volcanos <- lapply(1:length(res), function(i){
  plot(res[[i]][, 'logFC'], -log10(res[[i]][, 'padj']),
       pch = 20, col = 'gray', cex = .5, xlim = c(-10, 10), ylim = c(0, 15),
       ylab = '-log10(pvalue)', xlab = 'logFC',
       main = paste0('epsilon=',names(res)[i]))
  points(res[[i]][simDataIslam$indDE, 'logFC'], pch=1, col=3, cex=.5, lwd=1,
          -log10(res[[i]][simDataIslam$indDE, 'padj']))
})
par(mfrow = c(1, 1))

# par(mfrow = c(3, 3))
# volcanos <- lapply(1:length(resDESeq), function(i){
#   plot(resDESeq[[i]][, 'logFC'], -log10(resDESeq[[i]][, 'padj']),
#        pch = 20, col = 'gray', cex = .5, xlim = c(-10, 10), ylim = c(0, 15),
#        ylab = '-log10(pvalue)', xlab = 'logFC',
#        main = paste0('epsilon=',names(resDESeq)[i]))
#   points(resDESeq[[i]][simDataIslam$indDE, 'logFC'], pch=1, col=3, cex=.5, lwd=1,
#           -log10(resDESeq[[i]][simDataIslam$indDE, 'padj']))
# })
# par(mfrow = c(1, 1))

```


## nDE, TPR, FDR at pval=0.05
```{r nb}
#edgeR
tab_de = lapply(res, function(y){
  nDE = sum(y$padj <= 0.05, na.rm = TRUE)
  TPR = mean(simData$indDE %in% which( y$padj <= 0.05))
  FPR = mean(which(y$padj <= 0.05) %in% simData$indNonDE)
  c(nDE = nDE, TPR = TPR, FPR = FPR)
})
do.call(rbind, tab_de)

# #DESeq2
# tab_de = lapply(resDESeq, function(y){
#   nDE = sum(y$padj <= 0.05, na.rm = TRUE)
#   TPR = mean(simData$indDE %in% which( y$padj <= 0.05))
#   FPR = mean(which(y$padj <= 0.05) %in% simData$indNonDE)
#   c(nDE = nDE, TPR = TPR, FPR = FPR)
# })
# do.call(rbind, tab_de)
```

## TPR vs FDR
```{r plot}
trueDE = rep(0, nTags)
trueDE[simData$indDE] = 1
pp = COBRAData(pval = as.data.frame(do.call(cbind, lapply(res, '[[', 1))),
               padj = as.data.frame(do.call(cbind, lapply(res, '[[', 2))),
               truth = data.frame(status = trueDE))
cobraperf <- calculate_performance(pp, binary_truth = "status", thrs = 0.05)
cobraplot <- prepare_data_for_plot(cobraperf)
#png("~/Dropbox/phdKoen/singleCell/zinbwaveZinger/plots2/islamEpsEval.png", width=7,height=8, units="in", res=300)
plot_fdrtprcurve(cobraplot, plottype = c("curve", "points"), pointsize = 1,linewidth = .5) + xlab("FDP")
#dev.off()
#
# pp = COBRAData(pval = as.data.frame(do.call(cbind, lapply(resDESeq, '[[', 1))),
#                padj = as.data.frame(do.call(cbind, lapply(resDESeq, '[[', 2))),
#                truth = data.frame(status = trueDE))
# cobraperf <- calculate_performance(pp, binary_truth = "status", thrs = 0.05)
# cobraplotDESeq <- prepare_data_for_plot(cobraperf)
# plot_fdrtprcurve(cobraplot, plottype = c("curve", "points"), pointsize = 1,linewidth = .5) + xlab("FDP")
```

```{r}
plot_fdrtprcurve(cobraplot, plottype = c("curve", "points"), pointsize = .1, xaxisrange = c(0, 0.2),linewidth = .5) + xlab("FDP")

plot_fdrtprcurve(cobraplotDESeq, plottype = c("curve", "points"), pointsize = .1, xaxisrange = c(0, 0.2),linewidth = .5) + xlab("FDP")

```
