---
title: "Case study on neuronal cells - FPR on mocks"
author: "Fanny Perraudeau"
date: "`r Sys.Date()`"
output: 
  html_document: 
    fig_height: 7
    fig_width: 7
    toc: yes
    code_folding: hide
    toc_float: yes
---

```{r options, echo=FALSE, results="hide",mesasge=FALSE, error=FALSE, include=FALSE, autodep=TRUE}
knitr::opts_chunk$set(fig.align="center", cache=TRUE, error=FALSE, message=FALSE, warning=TRUE)
library(DESeq2)
library(edgeR)
library(limma)
library(zingeR)
library(zinbwave)

library(genefilter)

library(BiocParallel)
library(doParallel)
library(Biobase)
library(ggplot2)
```

```{r}
NCORES <- 2
registerDoParallel(NCORES)
register(DoparParam())
```

The goal of this document is to compare the FPR between zinbwave-DE, zingeR-DE, and other DE methods on mock communities. I'll first try to reproduce Figure 5 from zingeR preprint.

# Data
```{r data}
path = '~/Documents/scRNAseq/gitrepo/zingeRPaper/singleCellPaper/case/'
load(paste0(path, 'esetUsoskin.RData'))
eset=eset[rowSums(exprs(eset)>0)>=20,]
exprs(eset) <- apply(exprs(eset),2,function(x) {storage.mode(x) <- 'integer'; x})

file = paste0(path, "fpr/subsetMatrixUsoskinFPR_randomCellTypes.txt")
subsets <- read.table(file)
```

# Functions

Code for rna-seq and zinger rna-seq functions are adapted from zingeRPaper-master/singleCellPaper/case/fpr/runScriptsUsoskin_pickingSession.R

```{r sourceEdgeR}
runEdgeR <- function(e) {
  library(edgeR)
  condition = pData(e)$condition
  pickingSession = pData(e)$pickingSession
  design <- model.matrix(~ condition + pickingSession)
  dgel <- DGEList(exprs(e))
  dgel <- edgeR::calcNormFactors(dgel)
  dgel=estimateDisp(dgel,design)
  edger.fit <- glmFit(dgel, design)
  edger.lrt <- glmLRT(edger.fit, coef="conditionB")
  pvals <- edger.lrt$table$PValue
  padj <- p.adjust(pvals,method="BH")
  padj[is.na(padj)] <- 1
  logfc <- edger.lrt$table$logFC
  list(pvals=pvals, padj=padj, logfc = logfc)
}

```

```{r deseq2}
runDESeq2_poscounts <- function(e) {
  library(DESeq2)
  dds <- DESeqDataSetFromMatrix(exprs(e), colData=DataFrame(pData(e)),
                                design=~ condition + pickingSession)
  dds <- estimateSizeFactors(dds,type="poscounts")
  dds <- estimateDispersions(dds)
  dds <- nbinomWaldTest(dds, betaPrior=TRUE, modelMatrixType="standard")
  res <- results(dds, name="condition_B_vs_A")
  pvals <- res$pvalue
  padj <- res$padj
  pvals[is.na(pvals)] <- 1
  padj[is.na(padj)] <- 1
  logfc <- res$log2FoldChange
  list(pvals=pvals, padj=padj, logfc = logfc)
}
```

```{r sourceEdgeRzingeR}
# code from zingeRPaper-master/singleCellPaper/case/fpr/runScriptsUsoskin_pickingSession.R
runEdgeREMLibSize=function(e){
  library(edgeR)
  library(genefilter)
  condition = pData(e)$condition
  pickingSession = pData(e)$pickingSession
  design <- model.matrix(~ condition + pickingSession)
  d <- DGEList(exprs(e))
  d <- edgeR::calcNormFactors(d)
  #not adding a design matrix models the zeroes with the library size automatically
  effLogLibSize = log(d$samples$lib.size*d$samples$norm.factors)
  pickingSession = pData(e)[,"Picking sessions"]
  designZI = model.matrix(~ effLogLibSize + pickingSession)
  zeroWeights = zeroWeightsLS(d$counts, design, verbose = FALSE)
  d$weights = zeroWeights
  d=estimateDisp(d,design)
  fit <- glmFit(d, design)
  lrt = glmWeightedF(fit, coef=2, independentFiltering = TRUE)
  list(pvals = lrt$table$PValue, padj = lrt$table$padjFilter,
       logfc = lrt$table$logFC)
}
```

```{r zingerdeseq2}
pvalueAdjustment_kvdb <- function(baseMean, filter, pValue,
                             theta, alpha=0.05, pAdjustMethod="BH") {
  # perform independent filtering
    if (missing(filter)) {
      filter <- baseMean
    }
    if (missing(theta)) {
      lowerQuantile <- mean(filter == 0)
      if (lowerQuantile < .95) upperQuantile <- .95 else upperQuantile <- 1
      theta <- seq(lowerQuantile, upperQuantile, length=50)
    }

    # do filtering using genefilter
    stopifnot(length(theta) > 1)
    filtPadj <- filtered_p(filter=filter, test=pValue,
                           theta=theta, method=pAdjustMethod)
    numRej  <- colSums(filtPadj < alpha, na.rm = TRUE)
    # prevent over-aggressive filtering when all genes are null,
    # by requiring the max number of rejections is above a fitted curve.
    # If the max number of rejection is not greater than 10, then don't
    # perform independent filtering at all.
    lo.fit <- lowess(numRej ~ theta, f=1/5)
    if (max(numRej) <= 10) {
      j <- 1
    } else {
      residual <- if (all(numRej==0)) {
        0
      } else {
        numRej[numRej > 0] - lo.fit$y[numRej > 0]
      }
      thresh <- max(lo.fit$y) - sqrt(mean(residual^2))
      j <- if (any(numRej > thresh)) {
        which(numRej > thresh)[1]
      } else {
        1
      }
    }
    padj <- filtPadj[, j, drop=TRUE]
    cutoffs <- quantile(filter, theta)
    filterThreshold <- cutoffs[j]
    filterNumRej <- data.frame(theta=theta, numRej=numRej)
    filterTheta <- theta[j]

    return(list(padj=padj, filterThreshold=filterThreshold, filterTheta=filterTheta, filterNumRej = filterNumRej, lo.fit=lo.fit, alpha=alpha))

}


runDESeq2Zero <- function(e){
  library(DESeq2) ; library(genefilter)
  condition = pData(e)$condition
  pickingSession = pData(e)$pickingSession
  dse <- DESeqDataSetFromMatrix(exprs(e), colData=DataFrame(pData(e)),
                                design=~ condition + pickingSession)
  dse <- estimateSizeFactors(dse, type="poscounts")
  effLogLibSize <- log(colSums(counts(dse))*(1/sizeFactors(dse)))
  designZI=model.matrix(~effLogLibSize + pickingSession)
  zeroWeights = zeroWeightsLS(counts(dse), verbose=FALSE,
                              design=model.matrix(~condition + pickingSession),
                              colData=colData(dse), normalization="DESeq2_pos",
                              designZI=designZI)
  assays(dse)[["weights"]] = zeroWeights
  dse <- estimateDispersions(dse)
  dse <- nbinomWaldTest(dse, betaPrior=TRUE, modelMatrixType="standard")
  res <- results(dse, name="condition_B_vs_A")
  baseMean=unname(rowMeans(sweep(counts(dse),2,1/sizeFactors(dse),FUN="*")))
  pvalDesZero = 2*(1-pt(abs(res$stat),df=rowSums(zeroWeights)-2))
  padjusted = pvalueAdjustment_kvdb(pValue=pvalDesZero,filter=baseMean,alpha=0.05)
  list(pvals=pvalDesZero,padj=padjusted$padj,logfc=res$log2FoldChange)
}
```

```{r zinbwaveEdgeR}
computeZinbwaveWeights <- function(zinb, counts){
  mu <- getMu(zinb)
  pi <- getPi(zinb)
  theta <- getTheta(zinb)
  theta <- matrix(rep(theta, each = ncol(counts)), ncol = nrow(counts))
  nb_part <- dnbinom(t(counts), size = theta, mu = mu)
  zinb_part <- pi * ( t(counts) == 0 ) + (1 - pi) *  nb_part
  zinbwg <- ( (1 - pi) * nb_part ) / zinb_part 
  t(zinbwg)
}

runZinbwaveEdgeR <- function(e){
  condition = pData(e)$condition
  pickingSession = pData(e)$pickingSession
  design <- model.matrix(~ condition + pickingSession)
  # compute zinbwave weights
  library(zinbwave)
  zinb <- zinbFit(exprs(e), X = design, epsilon = 1e8)
  weights <- computeZinbwaveWeights(zinb, exprs(e))
  # use zingeR-edgeR
  d <- DGEList(exprs(e))
  d <- edgeR::calcNormFactors(d)
  d$weights <- weights
  d=estimateDisp(d, design)
  fit=glmFit(d,design)
  lrt=glmWeightedF(fit,coef=2, independentFiltering = TRUE)
  pvals = lrt$table$PValue
  list(pvals = pvals, padj = lrt$table$padjFilter, logfc = lrt$table$logFC)
}
```

```{r zinbwavedeseq2}
runZinbwaveDESeq2 <- function(e){
  condition = pData(e)$condition
  pickingSession = pData(e)$pickingSession
  design <- model.matrix(~ condition + pickingSession)
  # compute zinbwave weights
  library(zinbwave)
  zinb <- zinbFit(exprs(e), X = design, epsilon = 1e8)
  weights <- computeZinbwaveWeights(zinb, exprs(e))
  # use zingeR-DESeq2
  dse <- DESeqDataSetFromMatrix(exprs(e), colData=DataFrame(pData(e)),
                                design=~ condition + pickingSession)
  dse <- estimateSizeFactors(dse, type="poscounts")
  dimnames(weights) = NULL
  assays(dse)[["weights"]] = weights
  dse <- estimateDispersions(dse)
  dse = nbinomWaldTest(dse, betaPrior=TRUE, useT=TRUE, df=rowSums(weights)-2)
  resultsNames(dse)
  res = results(dse, name="conditionB")
  list(pvals = res$pvalue, padj = res$padj, logfc = res$log2FoldChange)
}
```

```{r}
runMAST <- function(e){
  require(MAST)
  counts <- exprs(e)
  tpm <- counts*1e6/colSums(counts)
  tpm <- log2(tpm+1)
  sca <- FromMatrix(tpm,
                    cData=data.frame(group=pData(e)$condition,
                                     pickingSession=pData(e)$pickingSession))
  
  # Adaptive thresholding from MAST vignette
  freq_expressed <- 0.2
  thres <- thresholdSCRNACountMatrix(assay(sca), nbins = 10, min_per_bin = 50,
                                     conditions = pData(e)$condition)
  #par(mfrow=c(5,4))
  #plot(thres)
  assays(sca) <- list(thresh=thres$counts_threshold, tpm=assay(sca))
  expressed_genes <- freq(sca) > freq_expressed
  sca <- sca[expressed_genes,]

  ngeneson <- apply(exprs(e),2,function(x) mean(x>0))
  CD <- colData(sca)
  CD$ngeneson <- ngeneson
  CD$cngeneson <- CD$ngeneson-mean(ngeneson)
  colData(sca) <- CD
  ## differential expression
  fit <- zlm(~ cngeneson + group + pickingSession, sca = sca,
             method = "bayesglm", ebayes = TRUE)
  L <- matrix(0, nrow = ncol(coef(fit, "D")))
  rownames(L) <- colnames(coef(fit, "D"))
  L["groupB",] <- 1
  lrFit <- lrTest(fit, hypothesis = L)
  pval <- lrFit[, 'hurdle', 'Pr(>Chisq)']
  padj <- p.adjust(pval, method = "BH")
  list(pvals = pval, padj = padj, logfc = NA)
}
```

```{r limmavoom}
runVoom <- function(e) {
  library(limma)
  condition = pData(e)$condition
  pickingSession = pData(e)$pickingSession
  design <- model.matrix(~ condition + pickingSession)
  dgel <- DGEList(exprs(e))
  dgel <- edgeR::calcNormFactors(dgel)
  v <- voom(dgel,design,plot=FALSE)
  fit <- lmFit(v,design)
  fit <- eBayes(fit)
  tt <- topTable(fit,coef="conditionB",n=nrow(dgel),sort.by="none")
  pvals <- tt$P.Value
  padj <- p.adjust(pvals,method="BH")
  padj[is.na(padj)] <- 1
  list(pvals=pvals, padj=padj, logfc=tt$logFC)
}
```

# Results

We remove genes with only zeros. It seems that edgeR and zingeR-edgeR don't need to remove these genes. What are the pvalues for these genes?

## Boxplots
```{r res}
algos <- list("DESeq2_poscounts"= runDESeq2_poscounts,
              "zingeR-DESeq2"   = runDESeq2Zero,
              "zinbwave-DESeq2" = runZinbwaveDESeq2,
              "edgeR"           = runEdgeR,
              "zingeR-edgeR"    = runEdgeREMLibSize,
              "zinbwave-edgeR"  = runZinbwaveEdgeR,
              "limma-voom"      = runVoom,
              "MAST"            = runMAST)
namesAlgos <- names(algos)
names(namesAlgos) <- namesAlgos
nreps <- 10
```

```{r,eval=FALSE}
res <- lapply(7:nreps, function(i) {
    cat(i," ")
    eLoop <- eset[,as.numeric(subsets[i,])]
    cat('Removing ', sum(rowSums(exprs(eLoop)) == 0), " genes with only 0's")
    eLoop <- eLoop[rowSums(exprs(eLoop)) != 0, ]
    condition=factor(rep(c("A","B"),each=45))
    pickingSession=factor(rep(rep(c("Cold","rt1","rt2"),each=15),2))
    pData(eLoop)$condition=condition
    pData(eLoop)$pickingSession=pickingSession
    resFPR <- lapply(namesAlgos, function(n){
      print(n)
      algos[[n]](eLoop)
    })
    save(resFPR, file = sprintf('FPR_.%srda', i))
    rm(resFPR)
    i
})
```

```{r}
res <- lapply(1:nreps, function(i) {
  load(sprintf('FPR_.%srda', i))
  resFPR
})
```


```{r mocksBoxplot}
hlp=lapply(res,function(replication){
  lapply(replication,function(method){
    pval=method$pvals
    pval[is.na(pval)]=1 #independent filtering
    mean(pval<=0.05)
  })
})

fprHat=Reduce(hlp,f=cbind)
fprHat=matrix(unlist(fprHat),nrow=length(algos),ncol=nreps,byrow=FALSE)
rownames(fprHat)=namesAlgos

boxplotData=data.frame(fpr=c(t(fprHat)),method=rep(namesAlgos,each=nreps))
ggplot(boxplotData,aes(x=reorder(method,fpr,median),y=fpr)) +
  geom_boxplot(outlier.colour=rgb(0,0,0,0)) + theme_bw() +
    geom_point(position = position_jitter(w = 0.1, h = 0),
               color="grey50", size=1) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) + xlab("") +
  scale_colour_discrete(guide="none")  + ylab("False Positive Rate") +
  geom_hline(aes(yintercept=0.05,colour="red"))
```

## Histograms of pvalues
```{r mocksPvalues}
par(mfrow=c(3,3))
hh <- sapply(1:length(res[[1]]), function(i){
  hist(res[[1]][[i]]$pvals, main=names(algos)[i], xlab="",
       breaks=seq(0,1,by=0.05)) 
})
par(mfrow=c(1,1))
```




